---
output: typst
editor_options: 
  chunk_output_type: console
---

# Estimation with 2X2 and staggered DiD

## Objectives

Le document présente l'estimation en double différence des groupes de traitement et de contrôle pour la période 2008 à 2021. Le principe est de comparer l'indice de richesse des ménages de contrôle et de traitement à une période antérieure et à une période postérieure à la mise en place des aires protégées. La période de pré-traitement sera la période de 1997 à 2008, et la période de post-traitement sera la période 2008 et 2021. Le groupe de traitement sont les ménages vivant à moins de 10km des aires protégées créées entre 2008 et 2021. Le groupe de contrôle sont les ménages vivant à plus de 10km des aires protégées créées jusqu'en 2021. Les ménages vivant à moins de 10km d'une aire protégée créée avant 2008 seront exclus de l'analyse.

Dans nos estimations par DID, nous avons utilisé des erreurs standards robustes aux corrélations intra-grappes. Cette estimation permet de gérer la dépendance des observations au sein des grappes. Le but est d'ajuster les erreurs standards en tenant compte des corrélations intra-grappes présentes au sein des groupes par l'échantillonnage de la variable de résultat @abadie2021. Cet ajustement se calcule sur la base des résidus du modèle et d'une estimation de variance intra-cluster. On calcule les résidus à partir du modèle de régression puis on utilise ces résidus pour construire une matrice variance-covariance robuste. La matrice va ajuster la corrélation intra-grappe en tenant compte la structure corporelle des erreurs.

Reminder, we have the following hypothesis:

-   H1-overall effect on livelihoods: PA reduce the living standards of nearby households by restricting access to natural resources, with insufficient compensation or benefits.

-   H2-effect on inequalities: PA exacerbate economic inequalities, as better-off or better-connected individuals capture most of the benefits (tourism jobs, development projects).

-   H3-heterogeneity: Impacts vary across PA depending on governance, community participation, and management approaches.

## Overall effect on livelihoods

### 2X2 DiD

In our PAP, we stated:

> "The difference-in-difference principle is to compare the wealth index of control and treatment households before and after the establishment of PA. Our treatment began in 2011, so we use 2008 DHS data for the pre-treatment and 2021 DHS data for the post-treatment. This method relies on the parallel trends assumption, meaning that, in the absence of treatment, treatment and control groups would have experienced similar changes over time. To verify this assumption, we include two pre-treatment, points (DHS 1997 and DHS 2008 data) and assess trends between these years. \[...\] The difference-in-difference method relies on the assumption of parallel trends. To validate this assumption, we will use as a reference, among the rural households surveyed in 1997, those living in an area located in or within of a PA created between 2008 and 2021 (placebo treatment group) and those matched to them using the method described above (placebo control group). We will graphically represent the evolution of the treatment and control groups in 1997, 2008 and 2021 for visual confirmation. To confirm the validity of the parallel trends assumption, we will conduct a placebo test 1997-2008 outcomes, as well as matching and control variables, while defining the treatment variable for post-2008 PA. If the estimated effect between 1997 and 2008 is null or statistically insignificant, this supports the parallel trends assumption."

```{r}
library(tidyverse)
library(haven)
library(fixest)
library(didimputation)
library(broom)
library(ggplot2)
library(gt)
library(gtExtras)
library(did2s)
library(modelsummary)

# Chargement des données
d97 <- read_rds("data/derived/data_matched_1997.rds") %>%
  rename(spei_wc_n_2 = spei_wc_1995,
         spei_wc_n_1 = spei_wc_1996,
         spei_wc_n   = spei_wc_1997) %>%
  mutate(hv219 = zap_labels(hv219), # hhh sex (1/2)
         hv220 = zap_labels(hv220)) # hhh age (num)

d08 <- read_rds("data/derived/data_matched_2008.rds") %>%
  rename(spei_wc_n_2 = spei_wc_2006,
         spei_wc_n_1 = spei_wc_2007,
         spei_wc_n   = spei_wc_2008) %>%
  mutate(hv219 = zap_labels(hv219),
         hv220 = zap_labels(hv220))

d11 <- read_rds("data/derived/data_matched_2011.rds") %>%
  rename(spei_wc_n_2 = spei_wc_2009,
         spei_wc_n_1 = spei_wc_2010,
         spei_wc_n   = spei_wc_2011) %>%
  mutate(hv219 = zap_labels(hv219),
         hv220 = zap_labels(hv220))

d13 <- read_rds("data/derived/data_matched_2013.rds") %>%
  rename(spei_wc_n_2 = spei_wc_2011,
         spei_wc_n_1 = spei_wc_2012,
         spei_wc_n   = spei_wc_2013) %>%
  mutate(hv219 = zap_labels(hv219),
         hv220 = zap_labels(hv220))

d16 <- read_rds("data/derived/data_matched_2016.rds") %>%
  rename(spei_wc_n_2 = spei_wc_2014,
         spei_wc_n_1 = spei_wc_2015,
         spei_wc_n   = spei_wc_2016) %>%
  mutate(hv219 = zap_labels(hv219),
         hv220 = zap_labels(hv220))

d18 <- read_rds("data/derived/data_matched_2018.rds") %>%
  rename(spei_wc_n_2 = spei_wc_2016,
         spei_wc_n_1 = spei_wc_2017,
         spei_wc_n   = spei_wc_2018) %>%
  mutate(hv219 = zap_labels(hv219),
         hv220 = zap_labels(hv220))

d21 <- read_rds("data/derived/data_matched_2021.rds") %>%
  rename(spei_wc_n_2 = spei_wc_2019,
         spei_wc_n_1 = spei_wc_2020,
         spei_wc_n   = spei_wc_2021) %>%
  mutate(hv219 = zap_labels(hv219),
         hv220 = zap_labels(hv220))

# Préparation des données
dat <- bind_rows(d97, d08, d11, d13, d16, d18, d21) %>%
  filter(GROUP %in% c("Treatment","Control")) %>%
  mutate(
    hv219   = factor(hv219, levels = c(1,2), labels = c("Homme","Femme")), # sexe (cat.)
    hv220   = as.numeric(hv220),                                           # âge
    treat   = as.integer(GROUP == "Treatment"),
    w_svy   = hv005 / 1e6,
    w_all   = w_svy * weights, # poids d'enquête × poids de matching (si 'weights' existe)
    id      = row_number(),
    treatment_phase = STATUS_YR,
    cluster_uid = interaction(DHSYEAR, hv001, drop = TRUE)
    )

# Outcome h
yvar <- "wealth_centile_rural_weighted"

# fixest DID 2×2: placebo 1997–2008, traitement 2008–2021 ------------------

# Placebo 1997–2008
pre <- dat %>%
  filter(DHSYEAR %in% c(1997, 2008)) %>%
  mutate(post = as.integer(DHSYEAR == 2008),
         treat_post = treat * post)

f_pre <- as.formula(paste(
  yvar, "~ treat + post + treat_post +",
  "spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220"
))

m_pre <- feols(f_pre, data = pre, weights = ~ w_all, cluster = ~ cluster_uid)

# Traitement 2008–2021
main <- dat %>%
  filter(DHSYEAR %in% c(2008, 2021)) %>%
  mutate(post = as.integer(DHSYEAR == 2021),
         treat_post = treat * post)

f_main <- f_pre  # même formule

m_main <- feols(f_main, data = main, weights = ~ w_all, cluster = ~ cluster_uid)

etable(m_pre, m_main, headers = c("Placebo 97–08", "Traitement 08–21"))

# Extraction compacte des deux effets DID
did_row <- function(model, year_post, vc = ~ cluster_uid, term = "treat_post"){
  summary(model, vcov = vc) %>% broom::tidy() %>%
    filter(term == !!term) %>%
    transmute(year = year_post, estimate, se = std.error)
}
did_df <- bind_rows(
  did_row(m_pre, 2008),
  did_row(m_main, 2021)
) %>%
  mutate(period = factor(ifelse(year == 2008, "1997–2008 (placebo)", "2008–2021 (effect)"),
                         levels = c("1997–2008 (placebo)", "2008–2021 (effect)")),
         lo = estimate - 1.96*se,
         hi = estimate + 1.96*se)

h1_2x2_plot <- ggplot(did_df, aes(x = period, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = .2) +
  geom_point(size = 3) +
  labs(x = NULL, y = "Effect on wealth index centile",
       title = "Impact on well-being: 2x2 DiD impact on local population (< 10km)")
       
write_rds(h1_2x2_plot, "soumission_articles/manuscript/figures/h1_2x2_plot.rds")
```

Le test placebo donne une estimation avec un coefficient faible (+1.8 centile) et non significatif: pas de différence détectée avant intervention dans les tendances du groupe traitement et du groupe contrôle. L'estimation du traitement présente une plus grande magnitude (-4.1 centiles), mais il reste non significatif. Les ménages vivant près des AP semblent connaître, un niveau de bien-être légèrement inférieur à celui des ménages contrôle, mais l'effet n'est pas significatif. A ce stade, l'analyse ne confirme pas l'hypothèse 1 (que les AP réduise le niveau de vie) : le sens de la relation est négatif, mais on ne peut pas rejeter l'hypothèse nulle. L'intervalle de confiance couvre encore largement 0, ce qui signifie qu'on a une précision insuffisante pour détecter un effet. L'effet est donc soit nul, soit inférieur à notre capacité de détection. Dans le PAP, on avait estimé que l'effet devrait être d'au moins 7.5 centiles pour pouvoir être détecté.

### Staggered diff-in-diff

#### Survey years and binning

In our PAP, we stated:

> "The MIS data can therefore be mobilized for this study. However, the periods covered by MIS only allow for comparisons beyond a simple “before” and “after” framework, which complicates the analysis and necessitates methods with limited methodological consensus. Since 2020, the state of the art in difference-in-difference methods has been challenged by the recognition that classical two-way fixed effects methods can yield spurious results when the effects of an intervention are heterogeneous and assessed over multiple periods. Since then, a dozen alternative approaches have been developed, but econometricians continue to debate their reliability (Roth et al. 2023). To avoid undermining the credibility of our results, we will initially limit ourselves to two study periods (DHS 2008 and 2021). Only if this approach lacks sufficient statistical power will we consider incorporating additional data. \[...\] Additionally, if we incorporate data from MIS 1997, MIS 2011, and MIS 2013, we could apply an estimator suitable for staggered adoption and multiple study periods accounting for potential heterogeneous treatment effects (Borusyak, Jaravel, and Spiess 2024). This will only be performed if power is insufficient with two periods of data."

We also add MICS 2018, as it presents the same design as MIS and DHS surveys, and has 17 870 surveyed households in a year for which we have no other data source. It enables us to have a more complete coverage of pre-and post year.

```{r}
library(tinytable)

# compute relative years and binned relative years
dat2 <- dat %>%
  mutate(
    treat_on = as.integer(treatment_phase > 0 & DHSYEAR >= treatment_phase),
    rel_year = if_else(treatment_phase > 0, DHSYEAR - treatment_phase, Inf),
    rel_year_clipped = case_when(
      is.infinite(rel_year) ~ NA_real_,
      TRUE ~ pmax(pmin(rel_year, 10), -10)
    ),
    rel_year_2yr = case_when(
      is.na(rel_year_clipped) ~ NA_integer_,
      rel_year_clipped <= -9  ~ -5,
      rel_year_clipped <= -7  ~ -4,
      rel_year_clipped <= -5  ~ -3,
      rel_year_clipped <= -3  ~ -2,
      rel_year_clipped <= -1  ~ -1,
      rel_year_clipped <=  1  ~  0,
      rel_year_clipped <=  3  ~  1,
      rel_year_clipped <=  5  ~  2,
      rel_year_clipped <=  7  ~  3,
      rel_year_clipped <=  9  ~  4,
      TRUE                    ~  5
    ),
    GROUP2 = case_when(GROUP == "Treatment" ~ "Treat.",
                      GROUP == "Control"   ~ "Ctrl.")
  ) 

## Label for binned years
bin_label <- function(k) {
  dplyr::case_when(
    is.na(k) ~ NA_character_,
    k == -5L ~ "min to -9",
    k == -4L ~ "-8 to -7",
    k == -3L ~ "-6 to -5",
    k == -2L ~ "-4 to -3",
    k == -1L ~ "-2 to -1",
    k ==  0L ~ "0 to 1",
    k ==  1L ~ "2 to 3",
    k ==  2L ~ "4 to 5",
    k ==  3L ~ "6 to 7",
    k ==  4L ~ "8 to 9",
    k ==  5L ~ "10 to max",
    TRUE     ~ as.character(k)
  )
}


# Build the combined table
tab_rel <- dat2 %>%
  mutate(cluster_uid = interaction(DHSYEAR, cluster_uid, drop = TRUE)) %>%
  group_by(rel_year, rel_year_2yr, GROUP2) %>%
  summarise(
    n_clusters_rel = n_distinct(cluster_uid),
    n_obs_rel = n(),
    .groups = "drop"
  )

tab_bin <- dat2 %>%
  mutate(cluster_uid = interaction(DHSYEAR, cluster_uid, drop = TRUE)) %>%
  group_by(rel_year_2yr, GROUP2) %>%
  summarise(
    n_clusters_bin = n_distinct(cluster_uid),
    n_obs_bin = n(),
    .groups = "drop"
  )

tab_both <- tab_rel %>%
  left_join(tab_bin, by = c("rel_year_2yr", "GROUP2")) %>%
  arrange(rel_year_2yr, GROUP2, rel_year) %>%
  group_by(rel_year_2yr, GROUP2) %>%
  mutate(
    block_n = n(),
    block_first = row_number() == 1,
    rel_year_2yr_lab = bin_label(rel_year_2yr),

    # show bin totals only once per block
    rel_year_2yr_show   = if_else(block_n > 1 & !block_first, "", rel_year_2yr_lab),
    n_clusters_bin_show = if_else(block_n > 1 & !block_first, NA_integer_, n_clusters_bin),
    n_obs_bin_show      = if_else(block_n > 1 & !block_first, NA_integer_, n_obs_bin)
  ) %>%
  ungroup() %>%
  select(
    rel_year,
    GROUP2,
    n_clusters_rel,
    n_obs_rel,
    rel_year_2yr = rel_year_2yr_show,
    n_clusters_bin = n_clusters_bin_show,
    n_obs_bin = n_obs_bin_show,
    block_n,
    block_first,
    rel_year_2yr_sort = rel_year_2yr
  )

tab_print <- tab_both %>% 
  select( rel_year, GROUP2, n_clusters_rel, n_obs_rel, rel_year_2yr, n_clusters_bin, 
  n_obs_bin, block_n, block_first, rel_year_2yr_sort )

gt_tab <- tab_print %>%
  select(-block_n, -block_first, -rel_year_2yr_sort) %>%
  gt() %>%
  cols_label(
    rel_year = "Rel. year",
    GROUP2 = "Group",
    n_clusters_rel = "Clusters",
    n_obs_rel = "Obs.",
    rel_year_2yr = "Bin",
    n_clusters_bin = "Clusters",
    n_obs_bin = "Obs."
  ) %>%
  fmt_number(columns = c(n_clusters_rel, n_obs_rel, n_clusters_bin, n_obs_bin), decimals = 0) %>%
  sub_missing(columns = c(n_clusters_bin, n_obs_bin), missing_text = "") %>%
  tab_options(
    table.font.size = px(10),
    data_row.padding = px(2),
    heading.padding = px(2),
    column_labels.padding = px(2)
  ) %>%
  cols_align(
    align = "left",
    columns = c(rel_year, GROUP2, rel_year_2yr)
  ) %>%
  cols_align(
    align = "right",
    columns = c(n_clusters_rel, n_obs_rel, n_clusters_bin, n_obs_bin)
  )
write_rds(gt_tab, "soumission_articles/table/rel_year_table.rds")
gt_tab
```

Pour l'instant, on n'arrive pas à appliquer l'approche de @borusyak2024 avec des données en coupes répétées (pas du panel) avec le package R didimputation. En théorie cette méthode peut s'appliquer à des données "repeated cross sections" et il semblerait qu'un package STATA le fasse. Mais nous ne disposons pas de STATA pour tester. On va donc appliquer la méthode de Gardner, qui en principe est très proche (avec des intervalles de confiance un peu plus large).

Nota bene : Afin de stabiliser les estimations de l’event-study, nous avons appliqué un binning des années relatives au traitement, regroupant toutes les observations au-delà de ±5 années dans des catégories communes. Ce choix est motivé par l’espacement irrégulier des enquêtes disponibles, qui implique que certaines valeurs de temps relatif sont représentées par très peu d’observations et produisent des coefficients instables. Comme le soulignent Borusyak, Jaravel et Spiess (2024, p. 22) : “*In practice, it is common to bin distant leads and lags into a single category, both to improve statistical precision and to avoid presenting very noisy and uninformative coefficients.*” Cette approche permet ainsi de réduire le bruit aux extrêmes et de se concentrer sur l’interprétation des dynamiques principales autour du traitement.s

## Estimation statique

```{r}
# Statique (traitement "on/off")
did2s_static <- did2s(
  data        = dat2,
  yname       = yvar,
  first_stage = ~ spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage= ~ i(treat_on, ref = FALSE),
  treatment   = "treat_on",
  cluster_var = "cluster_uid",
  weights     = "w_all"
)



h1_staggered_static_tb <- modelsummary(list("did2s (static)" = did2s_static))
write_rds(h1_staggered_static_tb, "soumission_articles/table/h1_staggered_static_tb.rds")
```

Now the event study

```{r}

dat2 <- dat2 |>
  mutate(
    rel_year_2yr2 = case_when(
      is.na(rel_year_2yr) ~ "never",
      TRUE ~ as.character(rel_year_2yr)
    ),
    rel_year_2yr2 = factor(
      rel_year_2yr2,
      levels = c("-5","-4","-3","-2","-1","0","1","2","3","4","5","never")
    )
  )

did2s_es_2yr <- did2s(
  data        = dat2,
  yname       = yvar,
  first_stage = ~ spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage= ~ i(rel_year_2yr2, ref = c("-1", "never")),
  treatment   = "treat_on",
  cluster_var = "hv001",
  weights     = "w_all"
)

etable(did2s_es_2yr, headers = "did2s event-study (2-yr bins ±10)")

# Tidy coefficients and plot
plot_did2s_2yr <- broom::tidy(did2s_es_2yr, conf.int = TRUE) %>%
  filter(grepl("^rel_year_2yr::", term)) %>%
  mutate(k = as.numeric(sub("^rel_year_2yr::", "", term))) %>%
  arrange(k)

plot_did2s_2yr <- broom::tidy(did2s_es_2yr, conf.int = TRUE) %>%
  filter(grepl("^rel_year_2yr2::", term)) %>%          # <- fix
  mutate(k = sub("^rel_year_2yr2::", "", term)) %>%    # keep as string (includes "never")
  mutate(k = factor(k, levels = c("-5","-4","-3","-2","0","1","2","3","4","5","never"))) %>%
  arrange(k)

h1_event_study_plot <- ggplot(plot_did2s_2yr, aes(k, estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = .2) +
  labs(
    x = "Year-to-treatment (binned by 2 and at ±10 years)",
    y = "Estimated effect on wealth centile",
    title = "Impact on well-being: Event-study with 2-year bins"
  ) + 
  scale_x_discrete(labels = bin_label)

write_rds(h1_event_study_plot, "soumission_articles/manuscript/figures/h1_event_study_plot.rds")

# Return relevant objects
list(
  did2s_tidy = plot_did2s_2yr,
  plot = h1_event_study_plot
)
```

## Effect on inequalities

### Estimation en 2x2 DiD

```{r}
# Outcome
yvar <- "zscore_wealth"

# fixest DID 2×2: placebo 1997–2008, traitement 2008–2021 ------------------

# Placebo 1997–2008
pre <- dat %>%
  filter(DHSYEAR %in% c(1997, 2008)) %>%
  mutate(post = as.integer(DHSYEAR == 2008),
         treat_post = treat * post)

f_pre <- as.formula(paste(
  yvar, "~ treat + post + treat_post +",
  "spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220"
))

m_pre <- feols(f_pre, data = pre, weights = ~ w_all, cluster = ~ cluster_uid)

# Traitement 2008–2021
main <- dat %>%
  filter(DHSYEAR %in% c(2008, 2021)) %>%
  mutate(post = as.integer(DHSYEAR == 2021),
         treat_post = treat * post)

f_main <- f_pre  # même formule

m_main <- feols(f_main, data = main, weights = ~ w_all, cluster = ~ cluster_uid)

etable(m_pre, m_main, headers = c("Placebo 97–08", "Traitement 08–21"))

# Extraction compacte des deux effets DID
did_row <- function(model, year_post, vc = ~ cluster_uid, term = "treat_post"){
  summary(model, vcov = vc) %>% broom::tidy() %>%
    filter(term == !!term) %>%
    transmute(year = year_post, estimate, se = std.error)
}
did_df <- bind_rows(
  did_row(m_pre, 2008),
  did_row(m_main, 2021)
) %>%
  mutate(period = factor(ifelse(year == 2008, "1997–2008 (placebo)", "2008–2021 (effect)"),
                         levels = c("1997–2008 (placebo)", "2008–2021 (effect)")),
         lo = estimate - 1.96*se,
         hi = estimate + 1.96*se)

h2_2x2_plot <- ggplot(did_df, aes(x = period, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = .2) +
  geom_point(size = 3) +
  labs(x = NULL, y = "Effect on z-score of wealth index centile",
       title = "Impact on inequalities: 2x2 DiD impact on local population (< 10km)")

write_rds(h2_2x2_plot, "soumission_articles/manuscript/figures/h2_2x2_plot.rds")
```

## Estimation statique

```{r}
# Statique (traitement "on/off")
did2s_static_h2 <- did2s(
  data        = dat2,
  yname       = yvar,
  first_stage = ~ spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage= ~ i(treat_on, ref = FALSE),
  treatment   = "treat_on",
  cluster_var = "cluster_uid",
  weights     = "w_all"
)

h2_staggered_static_tb <- modelsummary(list("did2s (static)" = did2s_static_h2))
write_rds(h2_staggered_static_tb, "soumission_articles/table/h2_staggered_static_tb.rds")
h2_staggered_static_tb
```

Now the event study

```{r}

dat2 <- dat2 |>
  mutate(
    rel_year_2yr2 = case_when(
      is.na(rel_year_2yr) ~ "never",
      TRUE ~ as.character(rel_year_2yr)
    ),
    rel_year_2yr2 = factor(
      rel_year_2yr2,
      levels = c("-5","-4","-3","-2","-1","0","1","2","3","4","5","never")
    )
  )

did2s_es_2yr <- did2s(
  data        = dat2,
  yname       = yvar,
  first_stage = ~ spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage= ~ i(rel_year_2yr2, ref = c("-1", "never")),
  treatment   = "treat_on",
  cluster_var = "hv001",
  weights     = "w_all"
)

etable(did2s_es_2yr, headers = "did2s event-study (2-yr bins ±10)")

# Tidy coefficients and plot
plot_did2s_2yr <- broom::tidy(did2s_es_2yr, conf.int = TRUE) %>%
  filter(grepl("^rel_year_2yr::", term)) %>%
  mutate(k = as.numeric(sub("^rel_year_2yr::", "", term))) %>%
  arrange(k)

plot_did2s_2yr <- broom::tidy(did2s_es_2yr, conf.int = TRUE) %>%
  filter(grepl("^rel_year_2yr2::", term)) %>%          # <- fix
  mutate(k = sub("^rel_year_2yr2::", "", term)) %>%    # keep as string (includes "never")
  mutate(k = factor(k, levels = c("-5","-4","-3","-2","0","1","2","3","4","5","never"))) %>%
  arrange(k)

h2_event_study_plot <- ggplot(plot_did2s_2yr, aes(k, estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = .2) +
  labs(
    x = "Year-to-treatment (binned by 2 and at ±10 years)",
    y = "Estimated effect on z-score of wealth centile",
    title = "Impact on inequalities: Event-study with 2-year bins"
  ) + 
  scale_x_discrete(labels = bin_label)

write_rds(h2_event_study_plot, "soumission_articles/manuscript/figures/h2_event_study_plot.rds")

# Return relevant objects
list(
  did2s_tidy = plot_did2s_2yr,
  plot = h2_event_study_plot
)
```

## Heterogeneity

L’étude des effets hétérogènes vise à déterminer quelles catégories de ménages bénéficient davantage – ou moins – de la création des aires protégées, et sous quelles conditions. Pour cela, nous analysons l'influence de la gouvernance en distinguant les aires protégées dites « strictes » (statuts IUCN I-IV) et les aires « multi-usages » (statuts IUCN V-VI).

### 2x2 DiD par type d’aire protégée

```{r}
yvar <- "wealth_centile_rural_weighted"

iucn_sets <- list(
  strict = c("Ia","Ib","II","III","IV"),
  multi  = c("V","VI")
)

is_strict <- function(x) x %in% iucn_sets$strict
is_multi  <- function(x) x %in% iucn_sets$multi

# DID 2x2-----------------------------------------------------
dat3 <- dat %>%
  mutate(
    treat_strict = as.integer(GROUP == "Treatment" & is_strict(IUCN_CAT)),
    treat_multi  = as.integer(GROUP == "Treatment" & is_multi(IUCN_CAT))
  )

# Placebo ---------------------------------------------------

# Strict

pre_strict <- dat3 %>%
  filter(
    DHSYEAR %in% c(1997, 2008),
    treat_strict == 1 | (GROUP == "Control")
  ) %>%
  mutate(post = as.integer(DHSYEAR == 2008))


f_pre_strict <- as.formula(paste(
  yvar, "~ treat_strict + post + treat_strict:post +",
  "spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220"
))

m_pre_strict <- feols(f_pre_strict, data = pre_strict, weights = ~ w_all, cluster = ~ hv001)

# Multi
pre_multi <- dat3 %>%
  filter(
    DHSYEAR %in% c(1997, 2008),
    treat_multi == 1 | (GROUP == "Control")
  ) %>%
  mutate(post = as.integer(DHSYEAR == 2008))


f_pre_multi <- as.formula(paste(
  yvar, "~ treat_multi + post + treat_multi:post +",
  "spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220"
))

m_pre_multi <- feols(f_pre_multi, data = pre_multi, weights = ~ w_all, cluster = ~ hv001)

# Main estimation ---------------------------

main_strict <- dat3 %>%
  filter(
    DHSYEAR %in% c(2008, 2021),
    treat_strict == 1 | (GROUP == "Control")
  ) %>%
  mutate(post = as.integer(DHSYEAR == 2021))

m_main_strict <- feols(f_pre_strict, data = main_strict, weights = ~ w_all, cluster = ~ hv001)

main_multi <- dat3 %>%
  filter(
    DHSYEAR %in% c(2008, 2021),
    treat_multi == 1 | (GROUP == "Control")
  ) %>%
  mutate(post = as.integer(DHSYEAR == 2021))

m_main_multi  <- feols(f_pre_multi, data = main_multi, weights = ~ w_all, cluster = ~ hv001)


did_row <- function(model, year_post, term){
  summary(model) %>% broom::tidy() %>%
    filter(term == !!term) %>%
    transmute(year = year_post, estimate, se = std.error)
}

did_df <- bind_rows(
  did_row(m_pre_strict, 2008, "treat_strict:post") %>% mutate(type = "strict"),
  did_row(m_main_strict, 2021, "treat_strict:post") %>% mutate(type = "strict"),
  did_row(m_pre_multi, 2008, "treat_multi:post") %>% mutate(type = "multi"),
  did_row(m_main_multi, 2021, "treat_multi:post") %>% mutate(type = "multi")
) %>%
  mutate(
    period = factor(ifelse(year == 2008, "1997–2008", "2008–2021"),
                    levels = c("1997–2008","2008–2021")),
    lo = estimate - 1.96*se,
    hi = estimate + 1.96*se
  )


pd <- position_dodge(width = 0.2)

h3_2x2_plot <- ggplot(did_df, aes(x = period, y = estimate, colour = type)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 3, position = pd) +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = .2, position = pd) +
  labs(x = NULL, y = "Effet DID (centile de richesse)",
       title = "DID 2×2 par type d’AP (strict vs usage multiple)")

write_rds(h3_2x2_plot, "soumission_articles/manuscript/figures/h3_2x2_plot.rds")

h3_2x2_plot 
```

### Estimation statique par type d’aire protégée

```{r}
# Strict only
d_strict <- dat2 %>%
  mutate(treat_on = as.integer(GROUP == "Treatment" & is_strict(IUCN_CAT))) %>%
  filter(treat_on == 1 | GROUP == "Control")

did2s_static_strict <- did2s(
  data         = d_strict,
  yname        = yvar,
  first_stage  = ~ spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(treat_on, ref = FALSE),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

# Multi only
d_multi <- dat2 %>%
  mutate(treat_on = as.integer(GROUP == "Treatment" & is_multi(IUCN_CAT))) %>%
  filter(treat_on == 1 | GROUP == "Control")

did2s_static_multi <- did2s(
  data         = d_multi,
  yname        = yvar,
  first_stage  = ~ spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(treat_on, ref = FALSE),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

h3_staggered_static_tb <- modelsummary(list(
  "Strict only" = did2s_static_strict,
  "Multi only"  = did2s_static_multi,
  "Both"        = did2s_static
  )
)
write_rds(h3_staggered_static_tb, "soumission_articles/table/h3_staggered_static_tb.rds")
h3_staggered_static_tb
```

Strict only shows a positive point estimate, about + 6.4 centiles, with a standard error of 3.4. This is marginally significant and suggests a weak positive association between proximity to strict PAs and the wealth index, relative to controls. Multi-purpose PAs show a small negative estimate, but not statistically distinguishable from zero.

### Estimation en event-study par type d’aire protégée

```{r}

dat2 <- dat2 |>
  mutate(
    rel_year_2yr2 = case_when(is.na(rel_year_2yr) ~ "never", TRUE ~ as.character(rel_year_2yr)),
    rel_year_2yr2 = factor(rel_year_2yr2, levels = c("-5","-4","-3","-2","-1","0","1","2","3","4","5","never"))
  )

# Baseline 
did2s_es_2yr_both <- did2s(
  data         = dat2,
  yname        = yvar,
  first_stage  = ~ spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_2yr2, ref = c("-1","never")),
  treatment    = "treat_on",
  cluster_var  = "hv001",
  weights      = "w_all"
)

# Strict only 
d_strict_es <- dat2 |>
  mutate(treat_on = as.integer(GROUP == "Treatment" & is_strict(IUCN_CAT))) %>%
  filter(treat_on == 1 | GROUP == "Control")

did2s_es_2yr_strict <- did2s(
  data         = d_strict_es,
  yname        = yvar,
  first_stage  = ~ spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_2yr2, ref = c("-1","never")),
  treatment    = "treat_on",
  cluster_var  = "hv001",
  weights      = "w_all"
)

# Multi only 
d_multi_es <- dat2 |>
  mutate(treat_on = as.integer(GROUP == "Treatment" & is_multi(IUCN_CAT))) |>
  filter(treat_on == 1 | GROUP == "Control")

did2s_es_2yr_multi <- did2s(
  data         = d_multi_es,
  yname        = yvar,
  first_stage  = ~ spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_2yr2, ref = c("-1","never")),
  treatment    = "treat_on",
  cluster_var  = "hv001",
  weights      = "w_all"
)

etable(
  did2s_es_2yr_strict,
  did2s_es_2yr_multi,
  did2s_es_2yr_both,
  headers = c("Strict only", "Multi only", "Both")
)

# Tidy + plot helper
tidy_es <- function(mod, type){
  broom::tidy(mod, conf.int = TRUE) |>
    filter(grepl("^rel_year_2yr2::", term)) |>
    mutate(
      k = sub("^rel_year_2yr2::", "", term),
      k = factor(k, levels = c("-5","-4","-3","-2","0","1","2","3","4","5","never")),
      type = type
    ) |>
    arrange(k)
}

es_all <- bind_rows(
  tidy_es(did2s_es_2yr_strict, "strict"),
  tidy_es(did2s_es_2yr_multi,  "multi"),
  tidy_es(did2s_es_2yr_both,   "both")
)

pd <- position_dodge(width = 0.3)

h3_event_study_plot <- ggplot(es_all, aes(k, estimate, colour = type)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(position = pd) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = .2, position = pd) +
  labs(
    x = "Year-to-treatment (2-year bins, ±10)",
    y = "Estimated effect on wealth centile",
    title = "Event-study (did2s): strict vs multi vs pooled"
  ) +
  scale_x_discrete(labels = bin_label)

write_rds(h3_event_study_plot, "soumission_articles/manuscript/figures/h3_event_study_plot.rds")


list(
  models = list(strict = did2s_es_2yr_strict, multi = did2s_es_2yr_multi, both = did2s_es_2yr_both),
  tidy   = es_all,
  plot   = h3_event_study_plot
)

```

The event study estimates are highly unstable once the sample is split by protected area type. At most relative time periods, coefficients are imprecisely estimated and vary substantially across horizons, with occasional statistically significant points that do not follow a coherent temporal pattern. This reflects the fact that, after separating strict and multi use protected areas and restricting the control group accordingly, the number of observations contributing to each type–cohort–period cell becomes small, especially at longer leads and lags. As a result, individual coefficients are driven by a limited set of cohorts and are sensitive to reweighting, making them difficult to interpret as meaningful dynamic effects. In this context, the event study mainly illustrates data limitations rather than well identified temporal responses, and does not support a stable or interpretable dynamic pattern by protected area type.
