---
output: typst
editor_options: 
  chunk_output_type: console
---

# Migration outcomes: household mobility

## Motivation

In the pre-analysis plan, we chose not to include migration among the control variables, as it is likely influenced both by the direct effects of protected areas and by broader changes in living standards in surrounding areas, making it a collider rather than a valid control. Instead, migration is treated as an outcome variable, since it is more plausibly a consequence of protected areas than an independent cause. This allows us to assess whether populations leave their territory because of protected areas, or whether migration is one of the mechanisms through which these areas affect living standard, without introducing bias into the analysis. In their paper, Jones et al. [-@Jones2018] state that human migration is often considered an important driver of land use change and a threat to protected area integrity. Understanding household mobility and behaviour at the forest frontier is important for guiding conservation and development policies.

## Variable definition

Among the three available data sources (DHS, MIS, MICS), the indicators that may help explain the movements of household members are as follows:

| Indicateur | Source | Level | Indicator definition |
|------------------|------------------|------------------|------------------|
| V104 | DHS/MIS (women) | Individual | Number of years the respondent has lived in the village, town, or city where she was interviewed. Visitors to the community are coded 96 |
| MV104 | DHS/MIS (men) | Individual | Number of years the respondent has lived in the village, town, or city where he was interviewed. Visitors to the community are coded 96 |
| WB15 | MICS (women) | Individual | MICS equivalent of V104 |
| MWB15 | MICS (men) | Individual | MICS equivalent of V104 |
| HV102 | DHS/MIS | Household | De jure resident status (usual resident) |
| HV103 | DHS/MIS | Household | De facto resident status (slept there the previous night) |
| V135 / MV135 | DHS/MIS | Individual | Whether the respondent is a usual resident or just a visitor to the household |
| V136 / MV136 | DHS/MIS | Household | Total number of household members (residents + visitors who slept there) |

## Data availability summary

We use the following indicators

| Year | Source   | V104 | MV104 | MWB15 | WB15 |
|------|----------|------|-------|-------|------|
| 1997 | DHS-III  | ✓    | ✓     | ✗     | ✗    |
| 2008 | DHS-V    | ✓    | ✓     | ✗     | ✗    |
| 2011 | MIS      | ✓    | ✓     | ✗     | ✗    |
| 2013 | MIS      | ✓    | ✓     | ✗     | ✗    |
| 2016 | MIS      | ✓    | ✓     | ✗     | ✗    |
| 2018 | MICS-6   | ✗    | ✗     | ✓     | ✓    |
| 2021 | DHS-VIII | ✓    | ✓     | ✗     | ✗    |

## Data loading and harmonization

```{r}

#| label: setup
#| message: false
#| warning: false


library(tidyverse)
library(haven)
library(fixest)
library(did2s)
library(broom)
library(ggplot2)
library(gt)
library(sf)
library(modelsummary)
library(labelled)


options(scipen = 999)
```

### Load matched household data

```{r}
#| label: load-matched

load_matched <- function(path, year, spei_years) {
  read_rds(path) |>
    st_drop_geometry() |>
    mutate(DHSYEAR = year) |>
    rename(
      spei_wc_n_2 = !!sym(paste0("spei_wc_", spei_years[1])),
      spei_wc_n_1 = !!sym(paste0("spei_wc_", spei_years[2])),
      spei_wc_n   = !!sym(paste0("spei_wc_", spei_years[3]))
    ) |>
    mutate(
      hv219 = zap_labels(hv219),
      hv220 = as.numeric(zap_labels(hv220))
    ) |>
    filter(GROUP %in% c("Treatment", "Control")) |>
    select(DHSYEAR, hv001, hv002, hv005, GROUP, WDPAID, STATUS_YR, IUCN_CAT,
           treatment, weights, spei_wc_n_2, spei_wc_n_1, spei_wc_n, hv219, hv220)
}

matched_all <- bind_rows(
  load_matched("data/derived/data_matched_1997.rds", 1997, c(1995, 1996, 1997)),
  load_matched("data/derived/data_matched_2008.rds", 2008, c(2006, 2007, 2008)),
  load_matched("data/derived/data_matched_2011.rds", 2011, c(2009, 2010, 2011)),
  load_matched("data/derived/data_matched_2013.rds", 2013, c(2011, 2012, 2013)),
  load_matched("data/derived/data_matched_2016.rds", 2016, c(2014, 2015, 2016)),
  load_matched("data/derived/data_matched_2018.rds", 2018, c(2016, 2017, 2018)),
  load_matched("data/derived/data_matched_2021.rds", 2021, c(2019, 2020, 2021))
)

# Vérification avant jointure 

matched_all |> count(DHSYEAR, GROUP)
matched_all |> summarise(prop_status_yr_na = mean(is.na(STATUS_YR)))
```


### Load migration's survey data

#### Définition du statut de résidence et de la migration 

```{r}
# Nettoyage de la durée de résidence (DHS/MIS: v104/mv104, MICS: wb15/mwb15)---
# 95 = toujours vécu ici | 96 = visiteur | 99 = ne sait pas

clean_duration <- function(x) {
  y <- as.numeric(x)
  y <- ifelse(y == 99, NA_real_, y)
  y
}

# Identification du statut de résidence 

derive_migrant <- function(duree){
  
  statut_residence <- case_when(
    duree == 95 ~ "toujours ici",
    duree == 96 ~ "visiteur",
    duree >= 0 & duree < 95 ~ "duree_connue",
    TRUE ~ NA_character_
  )
  
  # Indicateur de migration 
  migrant <- case_when(
    duree == 95 ~ OL,              # non migrant 
    duree == 96 ~ NA_integer_,     # visiteur
    duree >= 0 & duree < 95 ~ 1L,  # migrant
    TRUE ~ NA_integer_
  )
  
  # Indicateur de migration récente (migration depuis 5 ans ou moins)
  migrant_recent <- case_when(
    duree == 96 ~ NA_integer_,     # Visiteur
    duree == 95 ~ NA_integer_,     # non-migrant
    duree >= 0 & duree <= 5 ~ 1L,  # migrant récent
    duree > 5 & duree < 95 ~ 0L,    # migrant ancien
    TRUE ~ NA_integer_
  )
  
  tibble(
    statut_residence = statut_residence, 
    migrant = migrant, 
    migrant_recent = migrant_recent
    )
}

# Identification du type de migration 
derive_type_migration <- function(
    migrant,
    residence_precedente) {
  
  case_when(
    migrant == 1 & residence_precedente == 1 ~
      "urbain_vers_rural",
    
    migrant == 1 & residence_precedente == 2 ~
      "rural_vers_rural",
    
    migrant == 0 ~
      "non_migrant",
    TRUE ~ NA_character_
  )
}
```


##### DHS/MIS files (1997, 2008, 2011, 2013, 2016, 2021)

```{r}

# Helper DHS/MIS, Femmes-----------------------------

load_migration_ir <- function(path, year, source, extra_vars = c("v012", "v106", "v024"), filter_rural = TRUE){
  
  cols <- c("v001", "v002", "v005", "v021", "v023",  "v024", "v025", "v104", "v105", extra_vars)
  
  df <- read_dta(path, col_select = all_of(cols)) 
  
  if (filter_rural) {
    df <- df |> filter(v025 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      source_enquete = source,
      sexe = "femme",
      duree_residence = clean_duration(v104),
      residence_precedente = case_when(
        v104 %in% c(95,96) ~ NA_real_,
        TRUE ~ as.numeric(v105)
      ),
      # Année d'arrivée estimée dans le cluster actuel
      annee_arrivee = ifelse(!is.na(duree_residence)& duree_residence < 95,
                             DHSYEAR - duree_residence, NA_real_)
    )
  
  
  df <- bind_cols(df, derive_migrant(df$duree_residence))
  
  
  df <- df |>
    mutate(
      type_migration = case_when(
         migrant == 1 & residence_precedente == 1 ~ "urbain_vers_rural",
        migrant == 1 & residence_precedente == 2 ~ "rural_vers_rural",
        migrant == 0 ~ "non_migrant",
        TRUE ~ NA_character_
        
      )
    )
  
  df |> 
    select(v001, v002, v005, v021, v023, DHSYEAR, source_enquete, sexe, duree_residence, statut_residence, migrant, migrant_recent, residence_precedente, region_residence, all_of(extra_vars))
}


# Helper DHS/MIS, hommes ----------------------------------------------

load_migration_mr <- function(path, year, source, filtrer_rural = TRUE) {

  cols <- c("mv001", "mv002", "mv005", "mv021", "mv023", "mv104", "mv105", "mv025",
            "mv012", "mv106", "mv024")

  df <- read_dta(path, col_select = all_of(cols)) |>
    rename(
      v001 = mv001, v002 = mv002, v005 = mv005, v021 = mv021, v023 = mv023,
      v012 = mv012, v106 = mv106, v024 = mv024, v025 = mv025
    )

  if (filtrer_rural) {
    df <- df |> filter(v025 == 2)
  }

  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      source_enquete = source,
      sexe = "homme",
      duree_residence = clean_duration(mv104),
      residence_precedente = case_when(
        mv104 %in% c(95, 96) ~ NA_real_,
        TRUE ~ as.numeric(mv105)
      ),
      annee_arrivee = ifelse(!is.na(duree_residence) & duree_residence < 95,
                              DHSYEAR - duree_residence, NA_real_)
    )

  df <- bind_cols(df, derive_migrant(df$duree_residence))

  df <- df |>
    mutate(type_migration = derive_type_migration(migrant, residence_precedente))

  df |>
    select(v001, v002, v005, v021, v023, DHSYEAR, source_enquete, sexe,
           duree_residence, statut_residence, migrant, migrant_recent,
           residence_precedente, type_migration, annee_arrivee,
           v025, v012, v106, v024)
}

# Femme 
ir_1997 <- load_migration_ir("data/raw/dhs/DHS_1997/MDIR31DT/MDIR31FL.DTA", 1997, "DHS-III")
ir_2008 <- load_migration_ir("data/raw/dhs/DHS_2008/MDIR51DT/MDIR51FL.DTA", 2008, "DHS-V")
ir_2011 <- load_migration_ir("data/raw/dhs/MIS_2011/MDIR61DT/MDIR61FL.DTA", 2011, "MIS")
ir_2013 <- load_migration_ir("data/raw/dhs/MIS_2013/MDIR6ADT/MDIR6AFL.DTA", 2013, "MIS")
ir_2016 <- load_migration_ir("data/raw/dhs/MIS_2016/MDIR71DT/MDIR71FL.DTA", 2016, "MIS")
ir_2021 <- load_migration_ir("data/raw/dhs/DHS_2021/MDIR81DT/MDIR81FL.DTA", 2021, "DHS-VIII")

#Homme
mr_1997 <- load_migration_mr("data/raw/dhs/DHS_1997/MDMR31DT/MDMR31FL.DTA", 1997, "DHS-III")
mr_2008 <- load_migration_mr("data/raw/dhs/DHS_2008/MDMR51DT/MDMR51FL.DTA", 2008, "DHS-V")
mr_2011 <- load_migration_mr("data/raw/dhs/MIS_2011/MDMR61DT/MDMR61FL.DTA", 2011, "MIS")
mr_2013 <- load_migration_mr("data/raw/dhs/MIS_2013/MDMR6ADT/MDMR6AFL.DTA", 2013, "MIS")
mr_2016 <- load_migration_mr("data/raw/dhs/MIS_2016/MDMR71DT/MDMR71FL.DTA", 2016, "MIS")
mr_2021 <- load_migration_mr("data/raw/dhs/DHS_2021/MDMR81DT/MDMR81FL.DTA", 2021, "DHS-VIII")


```


##### MICS Data (2018)

```{r}

# Femme
load_migration_mics_femme <- function(path, year = 2018, source = "MICS-6",
                                       filtrer_rural = TRUE) {

  df <- read_dta(
    path,
    col_select = c(HH1, wmweight, wb15, wb16, WB4, welevel, windex5, HH6, HH7)
  )

  if (filtrer_rural) {
    df <- df |> filter(HH6 == 2)
  }

  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      source_enquete = source,
      sexe = "femme",
      duree_residence = clean_duration(wb15),
      residence_precedente = case_when(
        wb15 %in% c(95, 96) ~ NA_real_,
        TRUE ~ as.numeric(wb16)
      ),
      annee_arrivee = ifelse(!is.na(duree_residence) & duree_residence < 95,
                              DHSYEAR - duree_residence, NA_real_)
    )

  df <- bind_cols(df, derive_migrant(df$duree_residence))

  df <- df |>
    mutate(type_migration = derive_type_migration(migrant, residence_precedente))

  df |>
    transmute(
      v001 = HH1, v002 = NA_real_, v005 = wmweight, v021 = HH1, v023 = NA_real_,
      DHSYEAR, source_enquete, sexe,
      duree_residence, statut_residence, migrant, migrant_recent,
      residence_precedente, type_migration, annee_arrivee,
      v025 = HH6, v012 = WB4, v106 = welevel, v024 = HH7
    )
}

wm_2018 <- load_migration_mics_femme("data/raw/mics/MICS6_2018/wm.dta")

# Homme 
load_migration_mics_homme <- function(path, year = 2018, source = "MICS-6",
                                       filtrer_rural = TRUE) {

  df <- read_dta(
    path,
    col_select = c(HH1, mnweight, mwb15, mwb16, MWB4, mnlevel, windex5, HH6, HH7)
  )

  if (filtrer_rural) {
    df <- df |> filter(HH6 == 2)
  }

  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      source_enquete = source,
      sexe = "homme",
      duree_residence = clean_duration(mwb15),
      residence_precedente = case_when(
        mwb15 %in% c(95, 96) ~ NA_real_,
        TRUE ~ as.numeric(mwb16)
      ),
      annee_arrivee = ifelse(!is.na(duree_residence) & duree_residence < 95,
                              DHSYEAR - duree_residence, NA_real_)
    )

  df <- bind_cols(df, derive_migrant(df$duree_residence))

  df <- df |>
    mutate(type_migration = derive_type_migration(migrant, residence_precedente))

  df |>
    transmute(
      v001 = HH1, v002 = NA_real_, v005 = mnweight, v021 = HH1, v023 = NA_real_,
      DHSYEAR, source_enquete, sexe,
      duree_residence, statut_residence, migrant, migrant_recent,
      residence_precedente, type_migration, annee_arrivee,
      v025 = HH6, v012 = MWB4, v106 = mnlevel, v024 = HH7
    )
}


mn_2018 <- load_migration_mics_homme("data/raw/mics/MICS6_2018/mn.dta")

```

##### Combiner les données en une seule base 

```{r}
dhs_pooled <- bind_rows(
  ir_1997, ir_2008, ir_2011, ir_2013, ir_2016, wm_2018, ir_2021,
  mr_1997, mr_2008, mr_2011, mr_2013, mr_2016, mn_2018, mr_2021
)

# Vérifications
dhs_pooled |> count(DHSYEAR, statut_residence)
dhs_pooled |> count(DHSYEAR, v025)   # doit valoir 2 partout (filtrage rural)
```


### Join with matched data

```{r}
dhs_pooled <- dhs_pooled |>
  left_join(
    matched_all,
    by = c("DHSYEAR" = "DHSYEAR", "v001" = "hv001", "v002" = "hv002")
  )

# Taux de non-appariement (individus dont le ménage n'a pas été matché)
dhs_pooled |> group_by(DHSYEAR) |> summarise(prop_non_matche = mean(is.na(GROUP)))
```


### Identification des clusters arrivées avant/ après création de l'AP 

```{r}
dhs_pooled <- dhs_pooled |>
  mutate(
    arrivee_vs_AP = case_when(
      migrant == 0 ~ "toujours_ici",
      migrant == 1 & is.na(STATUS_YR) ~ NA_character_,
      migrant == 1 & annee_arrivee < STATUS_YR ~ "arrive_avant_AP",
      migrant == 1 & annee_arrivee >= STATUS_YR ~ "arrive_apres_AP",
      TRUE ~ NA_character_
    )
  )

dhs_pooled |> count(DHSYEAR, GROUP, arrivee_vs_AP)

# Tendance de la migration selon Treatment vs Control
tab_migration_AP <- svyby(~migrant, ~DHSYEAR + GROUP, dhs_design_matched, svymean, na.rm = TRUE)
print(tab_migration_AP)

ggplot(tab_migration_AP, aes(x = DHSYEAR, y = migrant, color = GROUP)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = migrant - se, ymax = migrant + se), width = 0.3) +
  theme_minimal() +
  labs(x = "Année d'enquête", y = "Proportion de migrants",
       title = "Migration selon la proximité aux aires protégées (Treatment vs Control)")

# Modèle : la migration dépend-elle du groupe, dans le temps, en contrôlant pour le climat ?
modele_migration_AP <- svyglm(
  migrant ~ factor(DHSYEAR) * GROUP + sexe + hv219 + hv220 + spei_wc_n,
  design = dhs_design_matched,
  family = quasibinomial()
)
summary(modele_migration_AP)
```


### Descriptive statistics 

























