---
output: typst
editor_options: 
  chunk_output_type: console
---

# Migration outcomes: household mobility

## Motivation

In the pre-analysis plan, we chose not to include migration among the control variables, as it is likely influenced both by the direct effects of protected areas and by broader changes in living standards in surrounding areas, making it a collider rather than a valid control. Instead, migration is treated as an outcome variable, since it is more plausibly a consequence of protected areas than an independent cause. This allows us to assess whether populations leave their territory because of protected areas, or whether migration is one of the mechanisms through which these areas affect living standard, without introducing bias into the analysis. In their paper, Jones et al. [-@Jones2018] state that human migration is often considered an important driver of land use change and a threat to protected area integrity. Understanding household mobility and behaviour at the forest frontier is important for guiding conservation and development policies.

## Variable definition

Among the three available data sources (DHS, MIS, MICS), the indicators that may help explain the movements of household members are as follows:

| Indicateur | Source | Level | Indicator definition |
|------------------|------------------|------------------|------------------|
| V104 | DHS/MIS (women) | Individual | Number of years the respondent has lived in the village, town, or city where she was interviewed. Visitors to the community are coded 96 |
| MV104 | DHS/MIS (men) | Individual | Number of years the respondent has lived in the village, town, or city where he was interviewed. Visitors to the community are coded 96 |
| WB15 | MICS (women) | Individual | MICS equivalent of V104 |
| MWB15 | MICS (men) | Individual | MICS equivalent of V104 |
| HV102 | DHS/MIS | Household | De jure resident status (usual resident) |
| HV103 | DHS/MIS | Household | De facto resident status (slept there the previous night) |
| V135 / MV135 | DHS/MIS | Individual | Whether the respondent is a usual resident or just a visitor to the household |
| V136 / MV136 | DHS/MIS | Household | Total number of household members (residents + visitors who slept there) |

## Data availability summary

We use the following indicators

| Year | Source   | V104 | MV104 | WB15 | MWB15 |
|------|----------|------|-------|-------|------|
| 1997 | DHS-III  | ✓    | ✗     | ✗     | ✗    |
| 2008 | DHS-V    | ✓    | ✓     | ✗     | ✗    |
| 2011 | MIS      | ✓    | ✗     | ✗     | ✗    |
| 2013 | MIS      | ✓    | ✗     | ✗     | ✗    |
| 2016 | MIS      | ✓    | ✗     | ✗     | ✗    |
| 2018 | MICS-6   | ✗    | ✗     | ✓     | ✓    |
| 2021 | DHS-VIII | ✓    | ✓     | ✗     | ✗    |


## Data loading and harmonization

```{r}

#| label: setup
#| message: false
#| warning: false


library(tidyverse)
library(haven)
library(fixest)
library(survey)
library(did2s)
library(broom)
library(ggplot2)
library(gt)
library(sf)
library(modelsummary)
library(labelled)


options(scipen = 999)

options(survey.lonely.psu = "adjust")
```

## Load matched household data

```{r}
#| label: load-matched

load_matched <- function(path, year, spei_years) {
  read_rds(path) |>
    st_drop_geometry() |>
    mutate(DHSYEAR = year) |>
    rename(
      spei_wc_n_2 = !!sym(paste0("spei_wc_", spei_years[1])),
      spei_wc_n_1 = !!sym(paste0("spei_wc_", spei_years[2])),
      spei_wc_n   = !!sym(paste0("spei_wc_", spei_years[3]))
    ) |>
    mutate(
      hv219 = zap_labels(hv219),
      hv220 = as.numeric(zap_labels(hv220))
    ) |>
    filter(GROUP %in% c("Treatment", "Control")) |>
    select(DHSYEAR, hv001, hv002, hv005, GROUP, WDPAID, STATUS_YR, IUCN_CAT,
           treatment, weights, spei_wc_n_2, spei_wc_n_1, spei_wc_n, hv219, hv220)
}

matched_all <- bind_rows(
  load_matched("data/derived/data_matched_1997.rds", 1997, c(1995, 1996, 1997)),
  load_matched("data/derived/data_matched_2008.rds", 2008, c(2006, 2007, 2008)),
  load_matched("data/derived/data_matched_2011.rds", 2011, c(2009, 2010, 2011)),
  load_matched("data/derived/data_matched_2013.rds", 2013, c(2011, 2012, 2013)),
  load_matched("data/derived/data_matched_2016.rds", 2016, c(2014, 2015, 2016)),
  load_matched("data/derived/data_matched_2018.rds", 2018, c(2016, 2017, 2018)),
  load_matched("data/derived/data_matched_2021.rds", 2021, c(2019, 2020, 2021))
)

# Vérification avant jointure 

matched_all |> count(DHSYEAR, GROUP)
matched_all |> summarise(prop_status_yr_na = mean(is.na(STATUS_YR)))
```


### Load migration's survey data

#### Définition du statut de résidence et de la migration 

```{r}
# Nettoyage de la durée de résidence (DHS/MIS: v104/mv104, MICS: wb15/mwb15)---
# 95 = toujours vécu ici | 96 = visiteur | 99 = ne sait pas

clean_duration <- function(x) {
  y <- as.numeric(x)
  y <- ifelse(y == 99, NA_real_, y)
  y
}

# Identification du statut de résidence 

derive_migrant <- function(duree){
  
  statut_residence <- case_when(
    duree == 95 ~ "toujours ici",
    duree == 96 ~ "visiteur",
    duree >= 0 & duree < 95 ~ "duree_connue",
    TRUE ~ NA_character_
  )
  
  # Indicateur de migration 
  migrant <- case_when(
    duree == 95 ~ 0L,              # non migrant 
    duree == 96 ~ NA_integer_,     # visiteur
    duree >= 0 & duree < 95 ~ 1L,  # migrant
    TRUE ~ NA_integer_
  )
  
  # Indicateur de migration récente (migration depuis 5 ans ou moins)
  migrant_recent <- case_when(
    duree == 96 ~ NA_integer_,     # Visiteur
    duree == 95 ~ NA_integer_,     # non-migrant
    duree >= 0 & duree <= 5 ~ 1L,  # migrant récent
    duree > 5 & duree < 95 ~ 0L,    # migrant ancien
    TRUE ~ NA_integer_
  )
  
  tibble(
    statut_residence = statut_residence, 
    migrant = migrant, 
    migrant_recent = migrant_recent
    )
}

# Identification du type de migration 
derive_type_migration <- function(
    migrant,
    residence_precedente) {
  
  case_when(
    migrant == 1 & residence_precedente == 1 ~
      "urbain_vers_rural",
    
    migrant == 1 & residence_precedente == 2 ~
      "rural_vers_rural",
    
    migrant == 0 ~
      "non_migrant",
    TRUE ~ NA_character_
  )
}
```


##### DHS/MIS files (1997, 2008, 2011, 2013, 2016, 2021)



```{r}

# Helper DHS/MIS, Femmes-----------------------------
# v012/mv012: respondent's age
# v106/mv106: Highest education level attended
# v024/mv024: De facto region of residence 
# v025/mv025: De facto type of place of residence
# v001/mv001: cluster number
# v002/mv002: household number 
# v005/mv005: sample weight
# v021/mv025: Primary sample unit
# v023/mv023: Sample strate
# v104/mv104: Number of years the respondent has lived
# v105/mv105: Type of place of previous residence 

load_migration_ir <- function(path, year, source,
                               extra_vars = c("v012", "v106", "v024"),
                               filter_rural = TRUE) {

  cols <- c("v001", "v002", "v005", "v021", "v023", "v024", "v025", "v104", "v105", extra_vars)

  df <- read_dta(path, col_select = all_of(cols))

  if (filter_rural) {
    df <- df |> filter(v025 == 2)
  }

  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      source_enquete = source,
      sexe = "femme",
      duree_residence = clean_duration(v104),
      residence_precedente = case_when(
        v104 %in% c(95, 96) ~ NA_real_,
        TRUE ~ as.numeric(v105)
      ),
      annee_arrivee = ifelse(!is.na(duree_residence) & duree_residence < 95,
                              DHSYEAR - duree_residence, NA_real_)  # annee_arrivee: année d'enquête moins durée de résidence pour situer l'arrivée dans le temps
    )

  df <- bind_cols(df, derive_migrant(df$duree_residence))

  df <- df |>
    mutate(type_migration = derive_type_migration(migrant, residence_precedente))

  df |>
    select(v001, v002, v005, v021, v023, DHSYEAR, source_enquete, sexe,
           duree_residence, statut_residence, migrant, migrant_recent,
           residence_precedente, type_migration, annee_arrivee,
           v025, all_of(extra_vars))
}

ir_1997 <- load_migration_ir("data/raw/dhs/DHS_1997/MDIR31DT/MDIR31FL.DTA", 1997, "DHS-III")
ir_2008 <- load_migration_ir("data/raw/dhs/DHS_2008/MDIR51DT/MDIR51FL.DTA", 2008, "DHS-V")
ir_2011 <- load_migration_ir("data/raw/dhs/DHS_2011/MDIR61DT/MDIR61FL.DTA", 2011, "MIS")
ir_2013 <- load_migration_ir("data/raw/dhs/DHS_2013/MDIR6ADT/MDIR6AFL.DTA", 2013, "MIS")
ir_2016 <- load_migration_ir("data/raw/dhs/DHS_2016/MDIR71DT/MDIR71FL.DTA", 2016, "MIS")
ir_2021 <- load_migration_ir("data/raw/dhs/DHS_2021/MDIR81DT/MDIR81FL.DTA", 2021, "DHS-VIII")

```


```{r}
# Helper DHS/MIS, hommes ----------------------------------------------
# v012/mv012: respondent's age
# v106/mv106: Highest education level attended
# v024/mv024: De facto region of residence 
# v025/mv025: De facto type of place of residence
# v001/mv001: cluster number
# v002/mv002: household number 
# v005/mv005: sample weight
# v021/mv025: Primary sample unit
# v023/mv023: Sample strate
# v104/mv104: Number of years the respondent has lived
# v105/mv105: Type of place of previous residence 

load_migration_mr <- function(path, year, source, filtrer_rural = TRUE) {

  if (!file.exists(path)) {
    warning("Fichier MR introuvable pour ", year, " (", source, ") : ", path,
            " -- pas de module hommes (questionnaire individuel) pour cette enquête. ",
            "Le fichier PR, s'il existe, ne contient pas de variable équivalente ",
            "a la duree de residence (hv104/hv105 = sexe/age, pas migration).")
    return(NULL)
  }

  cols_voulues <- c("mv001", "mv002", "mv005", "mv021", "mv023",
                     "mv104", "mv105", "mv025", "mv012", "mv106", "mv024")

  noms_reels   <- names(read_dta(path, n_max = 0))
  noms_lower   <- tolower(noms_reels)
  idx          <- match(cols_voulues, noms_lower)
  cols_reelles <- noms_reels[idx]

  manquantes <- cols_voulues[is.na(idx)]
  if (length(manquantes) > 0) {
    warning("Variables manquantes dans le fichier MR ", year, " (", source, ") : ",
            paste(manquantes, collapse = ", "))
  }
  cols_reelles <- cols_reelles[!is.na(cols_reelles)]

  df <- read_dta(path, col_select = all_of(cols_reelles))
  names(df) <- tolower(names(df))

  essentielles <- c("mv001", "mv005", "mv104", "mv105", "mv025")
  if (!all(essentielles %in% names(df))) {
    warning("Variables essentielles absentes pour ", year, " (", source,
            ") -- impossible de deriver la migration. Fichier ignore.")
    return(NULL)
  }

  df <- df |>
    rename(
      v001 = mv001, v002 = mv002, v005 = mv005,
      v021 = mv021, v023 = mv023,
      v012 = mv012, v106 = mv106, v024 = mv024, v025 = mv025
    )

  if (filtrer_rural) {
    df <- df |> filter(v025 == 2)
  }

  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      source_enquete = source,
      sexe = "homme",
      duree_residence = clean_duration(mv104),
      residence_precedente = case_when(
        mv104 %in% c(95, 96) ~ NA_real_,
        TRUE ~ as.numeric(mv105)
      ),
      annee_arrivee = ifelse(!is.na(duree_residence) & duree_residence < 95,
                              DHSYEAR - duree_residence, NA_real_)
    )

  df <- bind_cols(df, derive_migrant(df$duree_residence))

  df <- df |>
    mutate(type_migration = derive_type_migration(migrant, residence_precedente))

  cols_finales <- c("v001", "v002", "v005", "v021", "v023", "DHSYEAR",
                     "source_enquete", "sexe", "duree_residence", "statut_residence",
                     "migrant", "migrant_recent", "residence_precedente",
                     "type_migration", "annee_arrivee", "v025", "v012", "v106", "v024")
  cols_presentes <- intersect(cols_finales, names(df))

  df |> select(all_of(cols_presentes))
}

#Homme
mr_1997 <- load_migration_mr("data/raw/dhs/DHS_1997/MDPR31DT/MDPR31FL.DTA", 1997, "DHS-III")
mr_2008 <- load_migration_mr("data/raw/dhs/DHS_2008/MDMR51DT/MDMR51FL.DTA", 2008, "DHS-V")
mr_2011 <- load_migration_mr("data/raw/dhs/DHS_2011/MDIR61DT/MDIR61FL.DTA", 2011, "MIS")
mr_2013 <- load_migration_mr("data/raw/dhs/DHS_2013/MDMR6ADT/MDMR6AFL.DTA", 2013, "MIS")
mr_2016 <- load_migration_mr("data/raw/dhs/DHS_2016/MDMR71DT/MDMR71FL.DTA", 2016, "MIS")
mr_2021 <- load_migration_mr("data/raw/dhs/DHS_2021/MDMR81DT/MDMR81FL.DTA", 2021, "DHS-VIII")

```


##### MICS Data (2018)

```{r}

# Femme--------------------------------------------------

# HH1: cluster number
# HH2: Household number 
# WB4: respondent's age 
# WB6A: Highest education level attended (Préscolaire = 0, Primaire = 1, Secondaire 1 = 2, Secondaire 2 = 3, Supérieur = 4)
# WB16A: lieu de vie (Grande ville = 1, Autre ville = 2, Milieu rural = 3, non reponse = 9)
# mnweight: sample weight

load_migration_mics_femme <- function(path, year = 2018, source = "MICS-6") {

  df <- read_sav(
    path,
    col_select = c(HH1, HH2, wmweight, WB4, WB6A, WB15, WB16A)
  )

  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      source_enquete = source,
      sexe = "femme",
      duree_residence = clean_duration(WB15),
      residence_precedente = case_when(
        WB15 %in% c(95, 96) ~ NA_real_,
        TRUE ~ as.numeric(WB16A)
      ),
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )

  df <- bind_cols(df, derive_migrant(df$duree_residence))

  df <- df |>
    mutate(type_migration = derive_type_migration(migrant, residence_precedente))

  df |>
    transmute(
      v001 = HH1,
      v002 = HH2,
      v005 = wmweight,
      v021 = HH1,
      v023 = NA_real_,
      DHSYEAR,
      source_enquete,
      sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      type_migration,
      annee_arrivee,
      v012 = WB4,
      v106 = WB6A
    )
}

wm_2018 <- load_migration_mics_femme("data/raw/mics/2018/SPSS datasets/wm.sav")

# Homme-----------------------------------------------------------

# HH1: cluster number
# HH2: Household number 
# MWB4: respondent's age 
# WB6: Highest education level attended (Primaire = 1, Secondaire 1 = 2, Secondaire 2 = 3, Supérieur = 4)
# mwelevel: education 
# MWB15: Number of years the respondent has lived
# MWB16A: Type of place of previous residence
# MWB17: Region of previous residence
# mnweight: sample weight


load_migration_mics_homme <- function(
    path,
    year = 2018,
    source = "MICS-6",
    filtrer_rural = TRUE
) {

  # -------------------------------------------------------
  # 1. Lecture des données
  # -------------------------------------------------------

  df <- read_sav(
    path,
    col_select = c(
      HH1,
      HH2,
      mnweight,
      MWB4,
      MWB6,
      mwelevel,
      MWB15,
      MWB16A,
      MWB17,
      HH6
    )
  )

  # -------------------------------------------------------
  # 2. Garder uniquement les hommes vivant en milieu rural
  # -------------------------------------------------------

  if (filtrer_rural) {
    df <- df |>
      filter(HH6 == 2)
  }

  # -------------------------------------------------------
  # 3. Création des variables de base
  # -------------------------------------------------------

  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      source_enquete = source,
      sexe = "homme",

      # Âge du répondant
      v012 = as.numeric(MWB4),

      # Niveau d'éducation
      v106 = as.numeric(mwelevel),

      # Durée de résidence dans le lieu actuel
      duree_residence = clean_duration(MWB15),

      # Type de lieu de résidence précédent
      type_milieu_precedent = case_when(
        MWB16A == 1 ~ "ville",
        MWB16A == 2 ~ "bourgade",
        MWB16A == 3 ~ "rural",
        MWB16A %in% c(5, 8) ~ NA_character_,
        TRUE ~ NA_character_
      ),

      # Région de résidence précédente
      region_precedente = case_when(
        MWB17 == 96 ~ "hors_pays",
        MWB17 == 8 ~ NA_character_,
        TRUE ~ as.character(MWB17)
      ),

      # Année approximative d'arrivée
      annee_arrivee = ifelse(
        !is.na(duree_residence) &
          duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )

  # -------------------------------------------------------
  # 4. Indicateurs de migration
  # -------------------------------------------------------

  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )

  # -------------------------------------------------------
  # 5. Type de migration
  # -------------------------------------------------------

  df <- df |>
    mutate(
      type_migration = case_when(

        migrant == 0 ~ "non_migrant",

        migrant == 1 &
          type_milieu_precedent == "ville" ~
          "ville_vers_rural",

        migrant == 1 &
          type_milieu_precedent == "bourgade" ~
          "bourgade_vers_rural",

        migrant == 1 &
          type_milieu_precedent == "rural" ~
          "rural_vers_rural",

        TRUE ~ NA_character_
      )
    )

  # -------------------------------------------------------
  # 6. Harmonisation avec DHS/MIS
  # -------------------------------------------------------

  df |>
    transmute(
      # Identifiants
      v001 = HH1,
      v002 = HH2,

      # Poids d'échantillonnage
      v005 = mnweight,

      # Identifiant du cluster
      v021 = HH1,

      # Variable non disponible / non utilisée
      v023 = NA_real_,

      # Informations sur l'enquête
      DHSYEAR,
      source_enquete,
      sexe,

      # Migration
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      annee_arrivee,

      # Origine migratoire
      type_milieu_precedent,
      region_precedente,
      type_migration,

      # Caractéristiques individuelles
      v012,
      v106,

      # Milieu de résidence
      v025 = HH6
    )
}


mn_2018 <- load_migration_mics_homme("data/raw/mics/2018/SPSS datasets/mn.sav")

```


##### Combiner les données en une seule base 



### Descriptive statistics 

























