---
output: typst
editor_options: 
  chunk_output_type: console
---

# Migration outcomes: household mobility

## Motivation

In the pre-analysis plan, we chose not to include migration among the control variables, as it is likely influenced both by the direct effects of protected areas and by broader changes in living standards in surrounding areas, making it a collider rather than a valid control. Instead, migration is treated as an outcome variable, since it is more plausibly a consequence of protected areas than an independent cause. This allows us to assess whether populations leave their territory because of protected areas, or whether migration is one of the mechanisms through which these areas affect living standard, without introducing bias into the analysis. In their paper, Jones et al. [-@Jones2018] state that human migration is often considered an important driver of land use change and a threat to protected area integrity. Understanding household mobility and behaviour at the forest frontier is important for guiding conservation and development policies.

## Variable definition

Among the three available data sources (DHS, MIS, MICS), the indicators that may help explain the movements of household members are as follows:

| Indicator | Source | Level | Indicator definition |
|------------------|------------------|------------------|------------------|
| V104 | DHS/MIS (women) | Individual | Number of years the respondent has lived in the village, town, or city where she was interviewed. Visitors to the community are coded 96 |
| V135 / MV135 | DHS/MIS | Individual | Whether the respondent is a usual resident or just a visitor to the household |
| V136 / MV136 | DHS/MIS | Household | Total number of household members (residents + visitors who slept there) |
| MV104 | DHS/MIS (men) | Individual | Number of years the respondent has lived in the village, town, or city where he was interviewed. Visitors to the community are coded 96 |
| WB15 | MICS (women) | Individual | MICS equivalent of V104 |
| MWB15 | MICS (men) | Individual | MICS equivalent of V104 |
| HV102 | DHS/MIS | Household | De jure resident status (usual resident) |
| HV103 | DHS/MIS | Household | De facto resident status (slept there the previous night) |

Migration is not direcly observed in the DHS, MIS, and MICS data. Instead, it is derived from respondents' reported duration of continuous residence in their current place of residence (V104/MV104 in DHS and WB15/MWB15 in MICS). Unlike the DHS and MICS, the 2011, 2013, and 2016 MIS surveys do not include the migration/length of residence module. Although the corresponding variable (v104) is present in the files, it was not filled in (the column contains only "NA" entries). These three waves are therefore excluded from the migration analysis, which is based on a panel limited to 4 waves (1997, 2008, 2018, 2021), compared to 7 waves that do not depend on this variable. 

Based on this information, we construct an indicator for migration (`migrant`) and an indicator for recent migration (`migrant_recent`), defined as having arrived in the current place of residence within the previous five years. 

This measure captures whether an individual has migrated to their current place of residence rather than an annual migration flow. Consequently, the DID estimates interpreted as changes in the probability of having migrated to the current place of residence, rather than changes in an annual migration rate. 


## Data availability summary

We use the following indicators

| Year | Source   | V104 | MV104 | WB15 | MWB15 |
|------|----------|------|-------|-------|------|
| 1997 | DHS-III  | ✓    | ✗     | ✗     | ✗    |
| 2008 | DHS-V    | ✓    | ✓     | ✗     | ✗    |
| 2011 | MIS      | ✗    | ✗     | ✗     | ✗    |
| 2013 | MIS      | ✗    | ✗     | ✗     | ✗    |
| 2016 | MIS      | ✗    | ✗     | ✗     | ✗    |
| 2018 | MICS-6   | ✗    | ✗     | ✓     | ✓    |
| 2021 | DHS-VIII | ✓    | ✓     | ✗     | ✗    |


The migration outcome is constructed from individual-level questions on duration of residence. Female respondents are available consistently across survey waves, whereas male respondents are not. Therefore, the main analysis uses female respondents to ensure comparability across years. 


## Data loading and harmonization

```{r}

#| label: setup
#| message: false
#| warning: false


library(tidyverse)
library(haven)
library(purrr)
library(fixest)
library(survey)
library(did2s)
library(broom)
library(ggplot2)
library(gt)
library(sf)
library(modelsummary)
library(labelled)


options(scipen = 999)

options(survey.lonely.psu = "adjust")
```

## Load matched household data

```{r}
#| label: load-matched

load_matched <- function(path, year, spei_years) {
  read_rds(path) |>
    st_drop_geometry() |>
    mutate(DHSYEAR = year) |>
    rename(
      spei_wc_n_2 = !!sym(paste0("spei_wc_", spei_years[1])),
      spei_wc_n_1 = !!sym(paste0("spei_wc_", spei_years[2])),
      spei_wc_n   = !!sym(paste0("spei_wc_", spei_years[3]))
    ) |>
    mutate(
      hv219 = zap_labels(hv219),
      hv220 = as.numeric(zap_labels(hv220))
    ) |>
    filter(GROUP %in% c("Treatment", "Control")) |>
    select(DHSYEAR, hv001, hv002, hv005, GROUP, WDPAID, STATUS_YR, IUCN_CAT,
           treatment, weights, spei_wc_n_2, spei_wc_n_1, spei_wc_n, hv219, hv220)
}

matched_all <- bind_rows(
  load_matched("data/derived/data_matched_1997.rds", 1997, c(1995, 1996, 1997)),
  load_matched("data/derived/data_matched_2008.rds", 2008, c(2006, 2007, 2008)),
  load_matched("data/derived/data_matched_2011.rds", 2011, c(2009, 2010, 2011)),
  load_matched("data/derived/data_matched_2013.rds", 2013, c(2011, 2012, 2013)),
  load_matched("data/derived/data_matched_2016.rds", 2016, c(2014, 2015, 2016)),
  load_matched("data/derived/data_matched_2018.rds", 2018, c(2016, 2017, 2018)),
  load_matched("data/derived/data_matched_2021.rds", 2021, c(2019, 2020, 2021))
)

# Vérification avant jointure 

matched_all |> count(DHSYEAR, GROUP)
matched_all |> summarise(prop_status_yr_na = mean(is.na(STATUS_YR)))
```


## Load migration's survey data

### Définition du statut de résidence et de la migration 

```{r}
# Nettoyage de la durée de résidence -----------------------------------
# DHS/MIS (v104/mv104, 1997/2008/2011/2013/2016/2021)
# 95 = toujours vécu ici | 96 = visiteur | 97 = Inconsistent | 98 = Don't know | 99 = ne sait pas
# MICS (wb15/mwb15, 2018)
# 99 = Non reponse 

clean_duration <- function(x) {
  y <- as.numeric(zap_labels(x))
  y <- ifelse(y %in% c(97, 98, 99), NA_real_, y)
  y
}

# Identification du statut de résidence 

derive_migrant <- function(duree){
  
  statut_residence <- case_when(
    duree == 95 ~ "toujours ici",
    duree == 96 ~ "visiteur",
    duree >= 0 & duree < 95 ~ "duree_connue",
    TRUE ~ NA_character_
  )
  
  # Indicateur de migration 
  migrant <- case_when(
    duree == 95 ~ 0L,              # non migrant 
    duree == 96 ~ NA_integer_,     # visiteur
    duree >= 0 & duree < 95 ~ 1L,  # migrant
    TRUE ~ NA_integer_
  )
  
  # Indicateur de migration récente (migration depuis 5 ans ou moins)
  migrant_recent <- case_when(
    duree == 96 ~ NA_integer_,     # Visiteur
    duree == 95 ~ NA_integer_,     # non-migrant
    duree >= 0 & duree <= 5 ~ 1L,  # migrant récent
    duree > 5 & duree < 95 ~ 0L,    # migrant ancien
    TRUE ~ NA_integer_
  )
  
  tibble(
    statut_residence = statut_residence, 
    migrant = migrant, 
    migrant_recent = migrant_recent
    )
}

# Identification du type de migration 
derive_type_migration <- function(
    migrant,
    residence_precedente) {
  
  case_when(
    migrant == 1 & residence_precedente == "urbain"   ~ "urbain_vers_rural",
    migrant == 1 & residence_precedente == "rural"    ~ "rural_vers_rural",
    migrant == 1 & residence_precedente == "etranger" ~ "etranger_vers_rural",
    migrant == 0 ~ "non_migrant",
    TRUE ~ NA_character_
  )
}
```


### Harmonization of the education variable (v106 vs WB6A/MWB6A)

The variables related to education in the three data sources are not categorized in the same way. In the DHS/MIS Data, variable v106 uses only 4 categories (0 = No education, 1 = Primary, 2 = Secondary, 3 = Higher), whereas in the MICS data, variable WB6A for women and MWB6A for men use more categories (0 = Preschool, 1 = Primary, 2 = Seconday 1, 3 = Secondary 2, 4 = Higher, 8 = Don't know, 9 = No response). We will therefore harmonize these variables into 5 categories (None, Preschool, Primary, Secondary, Higher). 

```{r}
# Fonction 

education_labels <- c("Aucune", "Prescolaire", "Primaire", "Secondaire", "Supérieur")

recode_education_dhs <- function(x) {
  x <- as.numeric(zap_labels(x))
  case_when(
    x == 0 ~ 0L,  # Aucune
    x == 1 ~ 2L,  # Primaire
    x == 2 ~ 3L,  # Secondaire
    x == 3 ~ 4L,  # Superieur
    TRUE   ~ NA_integer_
  )
}

recode_education_mics <- function(x){
  x <- as.numeric(zap_labels(x))
  case_when(
    x == 0        ~ 1L,          # prescolaire 
    x == 1        ~ 2L,          # primaire
    x %in% c(2,3) ~ 3L,          # secondaire 1 + 2 -> secondaire
    x == 4        ~ 4L,          # superieur
    x %in% c(8,9) ~ NA_integer_, # NSP / pas de reponse
    TRUE          ~ NA_integer_
  )
}

var_education <- function(code){
  factor(code, levels = 0:4, labels = education_labels)
}

```


### Harmonization of place of residence (v105/mv105 vs WB16A/MWB16A)

The variables related to respondent's place of residence in the three data sources are not categorized in the same way. In the DHS/MIS Data, variable v105/mv105 uses these categories (0 = capital, large city, 1 = Small city, 2 = Town, 3 = Countryside, 4 = Abroad), whereas in the MICS data, variable WB16A for women and MWB16A for men use more categories ( 1 = Large city, 2 = Other city, 3 = rural area, 9 = No response). We will therefore harmonize these variables into 5 categories (None, Preschool, Primary, Secondary, Higher). 

#### Women Data

```{r}

# DHS 

lieu_precedent_dhs <- function(v105){
  x <- as.numeric(zap_labels(v105))
  case_when(
       x %in% c(0, 1, 2) ~ "urbain",   # Capital/large city, Small city, Town
    x == 3            ~ "rural",    # Countryside
    x == 4            ~ "etranger", # Abroad
    TRUE              ~ NA_character_
  )
}

# MICS

lieu_precedent_mics <- function(WB16A){
    x <- as.numeric(zap_labels(WB16A))
    case_when(
      x %in% c(1, 2) ~ "urbain", # Grande ville, Autre ville
      x == 3         ~ "rural",  # Milieu rural
      x == 9         ~ NA_character_, # Non reponse
      TRUE           ~ NA_character_
    )
}

```


### DHS files (1997, 2008, 2021)

The sampling stratum variable is designated as "v022" for the year 1997, and "v023" for all other DHS/MIS waves (2008,2021). We rename the "sample stratum" variable to v023.  

#### Women Data

```{r}

# Helper DHS, Femmes-----------------------------
# v001: cluster number
# v002: household number 
# v005: sample weight
# v012: respondent's age
# v021: Primary sample unit
# v023: Sample strate
# v024: De facto region of residence 
# v025: De facto type of place of residence
# v104: Number of years the respondent has lived
# v105: Type of place of previous residence 
# v106: Highest education level attended


load_migration_ir <- function(path, year, strate_var_source,
                              extra_vars = c("v012", "v106"),
                              filter_rural = TRUE) {
  
  cols <- c(
    "v001", "v002", "v005", "v021", strate_var_source,
    "v024", "v025", "v026", "v104", "v105", extra_vars
  )
  
  df <- read_dta(
    path,
    col_select = all_of(unique(cols))
  )
  
  if (strate_var_source != "v023") {
    df <- df |> rename(v023 = !!sym(strate_var_source))
  }
  
  if (filter_rural) {
    df <- df |> filter(v025 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "femme",
      duree_residence = clean_duration(v104),
      
      residence_precedente = case_when(
        v104 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_dhs(v105)
      ),
      
      region_precedente = NA_real_,
      
      taille_localite = as.numeric(zap_labels(v026)),
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      education = var_education(
        recode_education_dhs(v106)
      )
    )
  
  df |>
    select(
      v001, v002, v005, v021, v023,
      DHSYEAR, sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025, v024,
      all_of(extra_vars)
    )
}

ir_1997 <- load_migration_ir("data/raw/dhs/DHS_1997/MDIR31DT/MDIR31FL.DTA", 1997, "v023")
ir_2008 <- load_migration_ir("data/raw/dhs/DHS_2008/MDIR51DT/MDIR51FL.DTA", 2008, "v023")
#ir_2011 <- load_migration_ir("data/raw/dhs/DHS_2011/MDIR61DT/MDIR61FL.DTA", 2011, "v023")
#ir_2013 <- load_migration_ir("data/raw/dhs/DHS_2013/MDIR6ADT/MDIR6AFL.DTA", 2013, "v023")
#ir_2016 <- load_migration_ir("data/raw/dhs/DHS_2016/MDIR71DT/MDIR71FL.DTA", 2016, "v023")
ir_2021 <- load_migration_ir("data/raw/dhs/DHS_2021/MDIR81DT/MDIR81FL.DTA", 2021, "v023")

```


#### Men Data

```{r}
# Helper DHS, hommes ----------------------------------------------
# mv012: respondent's age
# mv106: Highest education level attended
# mv024: De facto region of residence 
# mv025: De facto type of place of residence
# mv001: cluster number
# mv002: household number 
# mv005: sample weight
# mv025: Primary sample unit
# mv023: Sample strate
# mv104: Number of years the respondent has lived
# mv105: Type of place of previous residence 

load_migration_mr <- function(path, year, filtrer_rural = TRUE) {
  
  cols <- c(
    "mv001", "mv002", "mv005", "mv021", "mv023",
    "mv104", "mv105", "mv025", "mv026",
    "mv012", "mv106", "mv024"
  )
  
  df <- read_dta(
    path,
    col_select = all_of(cols)
  ) |>
    rename(
      v001 = mv001,
      v002 = mv002,
      v005 = mv005,
      v021 = mv021,
      v023 = mv023,
      v012 = mv012,
      v106 = mv106,
      v024 = mv024,
      v025 = mv025
    )
  
  if (filtrer_rural) {
    df <- df |> filter(v025 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "homme",
      duree_residence = clean_duration(mv104),
      residence_precedente = case_when(
        mv104 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_dhs(mv105)
      ),
      region_precedente = NA_real_,
      taille_localite = as.numeric(zap_labels(mv026)),
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      education = var_education(
        recode_education_dhs(v106)
      )
    )
  
  df |>
    select(
      v001, v002, v005, v021, v023,
      DHSYEAR, sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025, v012, v106, v024
    )
}

#Homme
#mr_1997 <- load_migration_mr("data/raw/dhs/DHS_1997/MDPR31DT/MDPR31FL.DTA", 1997, "DHS-III")
mr_2008 <- load_migration_mr("data/raw/dhs/DHS_2008/MDMR51DT/MDMR51FL.DTA", 2008)
#mr_2011 <- load_migration_mr("data/raw/dhs/DHS_2011/MDIR61DT/MDIR61FL.DTA", 2011, "MIS")
#mr_2013 <- load_migration_mr("data/raw/dhs/DHS_2013/MDMR6ADT/MDMR6AFL.DTA", 2013, "MIS")
#mr_2016 <- load_migration_mr("data/raw/dhs/DHS_2016/MDMR71DT/MDMR71FL.DTA", 2016, "MIS")
mr_2021 <- load_migration_mr("data/raw/dhs/DHS_2021/MDMR81DT/MDMR81FL.DTA", 2021)

```


### MICS Data (2018)

#### Women data 

```{r}

# Femme--------------------------------------------------

# HH1: cluster number
# HH2: household number 
# HH6: De facto type of place of residence
# HH7: De facto region of residence 
# WB4: respondent's age
# WB6A: Highest education level attended
# WB15: Number of years the respondent has lived
# WB16A: Type of place of previous residence 
# wmweight: sample weight
# PSU: Primary sample unit
# stratum: Sample strate


# MICS - Femmes -----------------------------------------------------------

load_migration_mics_femme <- function(path, year = 2018,
                                      filter_rural = TRUE) {
  
  df <- read_sav(
    path,
    col_select = c(
      "HH1", "HH2", "HH6", "HH7", "stratum", "wmweight",
      "WB4", "WB6A", "WB15", "WB16A", "WB17"
    )
  )
  
  if (filter_rural) {
    df <- df |>
      filter(HH6 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "femme",
      
      duree_residence = clean_duration(WB15),
      
      residence_precedente = case_when(
        WB15 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_mics(WB16A)
      ),
      
      region_precedente = as.numeric(zap_labels(WB17)),
      
      taille_localite = NA_real_,
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      
      education = var_education(
        recode_education_mics(WB6A)
      )
    )
  
  df |>
    transmute(
      v001 = HH1,
      v002 = as.numeric(HH2),
      v005 = wmweight,
      v021 = HH1,
      v023 = as.numeric(zap_labels(stratum)),
      DHSYEAR,
      sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025 = HH6,
      v012 = WB4,
      v024 = HH7
    )
}

wm_2018 <- load_migration_mics_femme("data/raw/mics/2018/SPSS datasets/wm.sav")

```


#### Men data

```{r}

# Homme-----------------------------------------------------------

# HH1: cluster number
# HH2: Household number 
# MWB4: respondent's age 
# WB6: Highest education level attended (Primaire = 1, Secondaire 1 = 2, Secondaire 2 = 3, Supérieur = 4)
# mwelevel: education 
# MWB15: Number of years the respondent has lived
# MWB16A: Type of place of previous residence
# MWB17: Region of previous residence
# mnweight: sample weight


# MICS - Hommes -----------------------------------------------------------

load_migration_mics_homme <- function(path, year = 2018,
                                      filter_rural = TRUE) {
  
  df <- read_sav(
    path,
    col_select = c(
      "HH1", "HH2", "HH6", "HH7", "stratum", "mnweight",
      "MWB4", "MWB6A", "MWB15", "MWB16A", "MWB17"
    )
  )
  
  if (filter_rural) {
    df <- df |>
      filter(HH6 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "homme",
      
      duree_residence = clean_duration(MWB15),
      
      residence_precedente = case_when(
        MWB15 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_mics(MWB16A)
      ),
      
      region_precedente = as.numeric(zap_labels(MWB17)),
      
      taille_localite = NA_real_,
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      
      education = var_education(
        recode_education_mics(MWB6A)
      )
    )
  
  df |>
    transmute(
      v001 = HH1,
      v002 = as.numeric(HH2),
      v005 = mnweight,
      v021 = HH1,
      v023 = as.numeric(zap_labels(stratum)),
      DHSYEAR,
      sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025 = HH6,
      v012 = MWB4,
      v024 = HH7
    )
}


mn_2018 <- load_migration_mics_homme("data/raw/mics/2018/SPSS datasets/mn.sav")

```


### Combine the data into a single dataset 

```{r}
# DHS----------------------------------------------------------------
dhs_only <- bind_rows(
  haven::zap_labels(ir_1997),
  haven::zap_labels(ir_2008),
  haven::zap_labels(ir_2021),
  haven::zap_labels(mr_2008),
  haven::zap_labels(mr_2021)
)

dhs_only %>% count(DHSYEAR)

dhs_only %>% count(statut_residence, migrant, migrant_recent)

dhs_femmes <- dhs_only %>%
  filter(sexe == "femme")

# MIS--------------------------------------------------------------
#mis_only <- bind_rows(
#  haven::zap_labels(ir_2011),
#  haven::zap_labels(ir_2013),
#  haven::zap_labels(ir_2016)
#)

# MICS--------------------------------------------------------------
mics_only <- bind_rows(wm_2018,mn_2018)

# Combiner tous les données 
migration_data <- bind_rows(
  haven::zap_labels(dhs_only),
  haven::zap_labels(mics_only)
)
 
migration_data|> count(DHSYEAR, sexe)
migration_data |> count(DHSYEAR, v025)

```

In total, for the analysis of migration, we included 57,568 individuals (men and women aged 15 to 49) from four survey waves. 


### Join with matched data

```{r}
migration_data <- migration_data |>
  mutate(
    DHSYEAR = as.integer(DHSYEAR),
    v001    = as.numeric(v001),
    v002    = as.numeric(v002)
  )
 
matched_all <- matched_all |>
  mutate(
    DHSYEAR = as.integer(DHSYEAR),
    hv001   = as.numeric(hv001),
    hv002   = as.numeric(hv002)
  )

migration_data <- migration_data |>
  left_join(matched_all, by = c("DHSYEAR" = "DHSYEAR", "v001" = "hv001", "v002" = "hv002"))

stopifnot(all(c("GROUP", "STATUS_YR", "IUCN_CAT", "hv219", "hv220") %in% names(migration_data)))

# Contrôle de taux de correspondance 
diagnostic_matching <- migration_data |>
  group_by(DHSYEAR) |>
  summarise(
    clusters_totaux    = n_distinct(v001),
    n_individus        = n(),
    prop_non_matche    = mean(is.na(GROUP)),
    .groups = "drop"
  ) |>
  left_join(
    matched_all |> group_by(DHSYEAR) |> summarise(clusters_apparies = n_distinct(hv001), .groups = "drop"),
    by = "DHSYEAR"
  ) |>
  mutate(ratio_clusters_apparies = clusters_apparies / clusters_totaux)
 
print(diagnostic_matching)

migration_data <- migration_data |>
  mutate(poids_brut = ifelse(DHSYEAR == 2018, v005, v005 / 1000000)) |>
  group_by(DHSYEAR, sexe) |>
  mutate(poids_norm = poids_brut / mean(poids_brut, na.rm = TRUE)) |>
  ungroup() |>
  mutate(
    w_all            = poids_norm * weights,
    cluster_uid      = paste(DHSYEAR, v001, sep = "_"),
    treat            = as.integer(GROUP == "Treatment"),
  )


# Base contenant que les individus appartenant à des observations matchées et dont le statut migratoire est connu
migration_data <- migration_data |>
  mutate(
    matched = !is.na(GROUP)
  )

dat_migrant <- migration_data |>
  filter(!is.na(GROUP), !is.na(migrant))
 
dat_migrant |> count(DHSYEAR, sexe, treat)


```

After the data merge, we observe that the 2018 MICS data show a significantly lower rate of association wit protected areas than the other three waves (2.2% of clusters, compared with 15-25% for 1997, 2008, and 2021), reducing the usable estimation sample for this wave to approximately 370 observations. 


## Descriptive statistics 

### Descriptive Statistics of Migration in DHS Data
```{r}
# Nombre d'observation par année et par sexe 
dat_migrant %>% count(DHSYEAR, sexe) %>%
  group_by(DHSYEAR) %>%
  mutate(
    pct = 100 * n / sum(n)
  ) %>%
  ungroup()
```

All respondents in the 1997 survey were women, as noted at the beginning of the analysis. Starting in 2008, the composition was much more balanced, but still predominantly female (about two-thirs women and one-third men). 

```{r}
# Nombre de migrant par année et par sexe 
dat_migrant %>%
  group_by(DHSYEAR, sexe) %>%
  summarise(
    n = n(),
    migrant_n = sum(migrant == 1, na.rm = TRUE),
    migrant_pct = mean(migrant == 1, na.rm = TRUE) * 100
  ) %>%
  ungroup()
```

The data show a significant increase in the number of respondents over time. The ratio of women to men aged 15 to 49 is relatively stable (~ 26 to 49% women, 16 to 20% men), except in 1997, when only women were surveyed. 


```{r}
# Nombre de migration récente par année 
dat_migrant %>%
  count(migrant_recent) %>%
  mutate(
    pct = 100 * n / sum(n)
  )

dat_migrant %>%
  group_by(DHSYEAR) %>%
  summarise(
    n = n(),
    migrant_recent_n = sum(migrant_recent == 1, na.rm = TRUE),
    migrant_recent_pct = mean(
      migrant_recent == 1,
      na.rm = TRUE
    ) * 100
  )

# Nombre de migration récente parmi les migrants 
dat_migrant %>%
  filter(migrant == 1) %>%
  group_by(DHSYEAR) %>%
  summarise(
    n_migrants = n(),
    recent_n = sum(migrant_recent == 1, na.rm = TRUE),
    recent_pct = mean(
      migrant_recent == 1,
      na.rm = TRUE
    ) * 100
  )
```

Between 2008 et 2021, approximately 50% of migrants associated with protected areas were recent migrants (49.9% in 2008, 49.6% in 2021), with a slight decline in 2018 (43.1%). This stability in the rate of recent migration indicates a steady and continuous migratory flow in areas near protected areas, where the migrant population is renewed at a constant rate: each year, approximately half of the migrants are recent arrivals. It should be noted that this analysis focuses solely on identified migrants (2,669 individuals, or 26.2% of the matched sample), with the remaining observations corresponding to non-migrants. These results reflect active and predictable mobility in the study areas. 


```{r}
# Vérification du statut_residence par année 
dat_migrant %>%
  count(statut_residence) %>%
  mutate(
    pct = 100 * n / sum(n)
  )

dat_migrant %>%
  count(DHSYEAR, statut_residence) %>%
  group_by(DHSYEAR) %>%
  mutate(
    pct = 100 * n / sum(n)
  ) %>%
  ungroup()
```

The majority of migrants linked to protected areas are permanent residents ("always here"): 76.2% in 2008, 75.8% in 2021, and 63.4% in 2018. Conversely, only 23.8% to 36.6% of migrants have a known and measurable length of residence, meaning they migrated at a specific time. This predominance of permanent residents (about 3 out of 4) indicates that most of the migrants surveyed do not remember or do not have precise information about their year of arrival. The exception in 2018, when 36.6% had a known duration of residence, may reflect a different sample or improved data collection. These results suggest that only a minority of migrants can provide precise information on their duration of residence, limiting the temporal analysis of migration. 
 

```{r}
# Vérification de la durée de résidence par année 
dat_migrant %>%
  filter(statut_residence == "duree_connue") %>%
  group_by(DHSYEAR) %>%
  summarise(
    n = n(),
    moyenne = mean(duree_residence, na.rm = TRUE),
    mediane = median(duree_residence, na.rm = TRUE),
    ecart_type = sd(duree_residence, na.rm = TRUE),
    minimum = min(duree_residence, na.rm = TRUE),
    maximum = max(duree_residence, na.rm = TRUE)
  ) %>%
  ungroup()
```

Migrants matched to protected areas have lived at their current place of residence for an average of 8 to 9 years (8.04 years in 2008, 9.14 years in 2018, 8.69 years in 2021), with a median of around 6 - 7 years. This stability in the average duration indicates that the migrant population does not change drastically between survey years. The high standard deviations (7-8 years) reveal significant variability: some migrants have just arrived (minimum = 0 years), while others have been settled for a long time (maximum = 40-51 years).


```{r}
# Vérification de l'année d'arrivée sur le lieu de résidence par année d'enquête 
dat_migrant %>%
  filter(!is.na(annee_arrivee)) %>%
  group_by(DHSYEAR) %>%
  summarise(
    moyenne = mean(annee_arrivee),
    mediane = median(annee_arrivee),
    minimum = min(annee_arrivee),
    maximum = max(annee_arrivee)
  ) %>%
  ungroup()
```

Among migrants with a known length of residence, the median year of arrival naturally increases over time: 2002 in 2008, 2011 in 2018, and 2015 in 2021. This simply reflects the fact that migrants interviewed in more recent surveys arrived later. On average, migrants surveyed in 2008 arrived around 2000, those surveyed in 2018 around 2009, and those surveyed in 2021 around 2012. The majority of arrivals occured between 200 and 2015, with a few very long-term migrants dating back to 1968-1983. These results confirm thata reas near protected areas experienced significant migration beginning in the 2000s, with a continuous influx through 2021. 



```{r}
# Type de migration par année 
dat_migrant %>%
  count(type_migration) %>%
  mutate(
    pct = 100 * n / sum(n)
  )

dat_migrant %>%
  count(DHSYEAR, type_migration) %>%
  group_by(DHSYEAR) %>%
  mutate(
    pct = 100 * n / sum(n)
  ) %>%
  ungroup()

dat_migrant %>% count(type_migration)

```

Migration in areas near protected areas is dominated by rural-to-rural migration: 17.5% in 2008, 31.6% in 2018, and 16.9% in 2021. Urban-to-rural migration is a minority but still present: 5.89% in 2008, 5.08% in 2018, and 7.25% in 2021. These results indicate that the majority of migrants come from other rural areas (17-32%), while a small proportion come from urban areas (5-7%). The year 2018 stands out with a higher share of rural-to-rural migration (31.6%) and a lower proportion of non-migrants (63.4%), suggesting a different survey profile. Overall, these data show that protected areas primarily attract migrants from other rural areas, confirming intra-regional rural mobility rather than an urban exodus. 



```{r}
# Analyse descriptive de migrants et non-migrants 
dat_migrant %>%
  group_by(migrant) %>%
  summarise(
    n = n(),
    age_mean = mean(v012, na.rm = TRUE),
    age_median = median(v012, na.rm = TRUE),
    education = mean(as.numeric(education), na.rm = TRUE),
    taille_localite_mean = mean(
      taille_localite,
      na.rm = TRUE
    )
  )

# Ajouter un test t
t.test(v012 ~ migrant, data = dat_migrant)
t.test(as.numeric(education) ~ migrant, data = dat_migrant)

```

Migrants and non-migrants share very similar characteristics. The average age of migrants (30.1 years) is slightly higher than that of non-migrants (29.0 years), with medians of 29 and 27 years, respectively. The educational level of migrants (2.88) is also slightly higher than that of non-migrants (2.58), suggesting that migrants, on average, possess slightly higher educational capital. However, these differences remain modest, indicating that migration to areas near protected areas does not select a very distinct population: migrants and non-migrants share comparable demographic and educational profiles. 



```{r}
#| label: setup-svy-design 
# Primary sampling unit =v021, Sample strate = v023, weights = w_all

dat_migrant_svy <- dat_migrant |>
  filter(!is.na(w_all), !is.na(v021), !is.na(v023), !is.na(GROUP))


design_migration <- svydesign(
  ids     = ~v021,
  strata  = ~v023,
  weights = ~w_all,
  data    = dat_migrant_svy,
  nest    = TRUE
)

```


```{r}
#| label: descriptive

# Statistiques pondérées (design-based) par année et par groupe des migrants récents
migrant_desc_svy <- svyby(
  ~migrant + migrant_recent,
  ~DHSYEAR + GROUP,
  design_migration,
  svymean,
  na.rm = TRUE,
  vartype = "ci"
) |>
  as_tibble() |>
  arrange(DHSYEAR, GROUP)

migrant_desc_svy

# Effectifs bruts 
dat_migrant_svy |> 
  summarise(n = n(), .by = c(DHSYEAR, GROUP))
```

Recent migration rates are similar between the Control and Treatment groups: approximately 50% in 2008 (52.8% for Control vs 48.6% for Treatment), 40% in 2018 (40.4% for Control and 40.1% for Treatment), and 50% in 2021 (52.3% for Control vs 49.2% for Treatment). This similarity in rates between the groups suggests no visible difference in the impact of protected areas on recent migration, at least at this stage of the analysis. 


```{r}
# Type de migration par année et groupe (pondéré)
type_dat_migrant_svy <- svytable(
  ~DHSYEAR + GROUP + type_migration,
  design_migration
) |>
  as_tibble() |>
  mutate(pct = n / sum(n) * 100, .by = c(DHSYEAR, GROUP))

type_dat_migrant_svy

dat_migrant |> group_by(DHSYEAR, sexe) |>
  summarise(pct_na_type_migration = mean(is.na(type_migration)), n = n())
```

Rural-to-rural migration predominates, but with a notable variation in 2018: it accounts for 53.3% in the Control Group versus 72.9% in the treatment group. In 2008 and 2021, the differences are minimal (17-18% for rural-to-rural, 5-8% for urban-to-rural). The year 2018 stands out with a very different profile: 53.29% of non-migrants in the Control group versus 72.9% in the Treatment group, suggesting a different survey profile. 


```{r}
#| label: fig-migration-rate
#| fig-cap: "Share of migrants by survey year and treatment group (weighted)"

ggplot(migrant_desc_svy, aes(x = factor(DHSYEAR), y = migrant * 100, fill = GROUP)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(
    aes(ymin = ci_l.migrant * 100, ymax = ci_u.migrant * 100),
    position = position_dodge(width = 0.7), width = 0.2
  ) +
  labs(x = "Survey year", y = "Share of migrants (%)", fill = "Group",
       title = "Migration prevalence by year and treatment status (weighted)") +
  theme_minimal(base_size = 13)
```



```{r}
#| label: tbl-balance
#| tbl-cap: "Balance of covariates between treatment and control groups"

dat_migrant_svy |>
  mutate(
    education = fct_recode(education,
      "No education" = "Aucune",
      "Pre-primary" = "Prescolaire",
      "Primary" = "primary",
      "Secondary" = "secondary",
      "Higher" = "higher"
    )
  ) |>
  rename(
    "Age" = v012,
    "Education" = education,
    "Sex" = hv219,
    "Household head age" = hv220,
    "Duration of residence" = duree_residence,
    "Locality size" = taille_localite
  ) |>
  select(GROUP, "Age", "Education", "Sex", 
         "Household head age", "Duration of residence", "Locality size") |>
  datasummary_balance(~GROUP, data = _, output = "gt")
```

The following balance table compares the sociodemographic characteristics between the treatment and control groups. In the control group, respondents are, on average, women aged 29, with a household head aged 40. Approximately 50.4% have a primary school education. The average length of residence in the area is 73 years. In the treatment group, respondents are, on average, mean aged 29, with a household head also aged 40. The educational distribution is similar to that of the control group. As for the population's length of residence, it averages 71.1%. 


```{r}
#| label: fig-migration-recent
#| fig-cap: "Share of recent migrants by survey year and treatment group (weighted)"

ggplot(migrant_desc_svy, aes(x = factor(DHSYEAR), y = migrant_recent * 100, fill = GROUP)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(
    aes(ymin = ci_l.migrant_recent * 100, ymax = ci_u.migrant_recent * 100),
    position = position_dodge(width = 0.7), width = 0.2
  ) +
  labs(x = "Survey year",
       y = "Share of recent migrants (%)",
       fill = "Group",
       title = "Recent migration prevalence by year and treatment status (weighted)") +
  theme_minimal(base_size = 13)
```



```{r}
#| label: fig-migration-type
#| fig-cap: "Distribution of migration types by survey year and treatment group (weighted)"

ggplot(type_dat_migrant_svy, aes(x = factor(DHSYEAR), y = pct, fill = type_migration)) +
  geom_col(position = "stack", width = 0.6) +
  facet_wrap(~GROUP) +
  labs(x = "Survey year",
       y = "Share (%)",
       fill = "Type de migration",
       title = "Composition des types de migration par année et groupe (pondéré)") +
  theme_minimal(base_size = 13)
```



## Migration analysis

### 2×2 DiD

We follow the same 2×2 DiD design as the main analysis: a placebo test (1997–2008) and the treatment estimation (2008–2021). The model is:

$$Migrant_{ict} = \alpha + \beta_1 \text{Treat}_c + \beta_2 \text{Post}_t + \delta (\text{Treat}_c \times \text{Post}_t) + \mathbf{X}_{ct}'\gamma + \varepsilon_{ict}$$

where $i$ indexes individuals (women or men depending on the specification), $c$ clusters (PSU), $t$ survey rounds. $Migrant_{ict}$ is a binary variable (moved to current place of residence vs. has always lived there). Controls $\mathbf{X}$ include SPEI values, household head sex and age, education level and the individual's gender. Standard errors are clustered at the cluster-year level and all estimates are weighted by the combined survey weights $w_{ict}$. 


```{r}
#| label: migrant-2x2

# Placebo test: 1997–2008 (no treatment expected)
pre_migrant <- dat_migrant |>
  filter(DHSYEAR %in% c(1997, 2008)) |>
  mutate(
    post = as.integer(DHSYEAR == 2008),
    treat_post = treat * post
  )

migrant_pre <- feols(
  migrant ~ treat + post + treat_post + sexe + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = pre_migrant,
  weights = ~w_all,
  cluster = ~cluster_uid
)


# Main analysis: 2008–2021 (treatment period)
main_migrant <- dat_migrant |>
  filter(DHSYEAR %in% c(2008, 2021)) |>
  mutate(
    post = as.integer(DHSYEAR == 2021),
    treat_post = treat * post
  )

migrant_main <- feols(
  migrant ~ treat + post + treat_post + sexe + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = main_migrant,
  weights = ~w_all,
  cluster = ~cluster_uid
)


etable(
  migrant_pre,
  migrant_main,
  headers = c(
    "Placebo 1997-2008\n(Full sample)",
    "Treatment 2008-2021\n(Full sample)"
  ),
  title = "DID Estimates: Migration Effects (Both sexes)"
)
```

The results show that there is no statistically significant effect of the treatment on the probability of migration. The next figure presents the results of the 2x2 DID estimation comparing the treatment groups (households living within 10 km of protected area) and the control groups (household located more than 10 Km away) before (1997-2008) and after (2008-2021) the creation of the protected areas. The placebo test for the pre-intervention period (1997-2008) yields a small ans statistically insignificant coefficient (-0.0265), indicating that no differential trend was detected prior to the intervention. This confirms the parallel trends hypothesis. In contrast, for the intervention period (2008-2021), the estimated effect is positive but statistically insignificant (+5.23). Beyond the treatment effect, several factors are strongly associated with migration: women living in male-headed households are 14% less likely to migrate, while higher education increases the probability of migration by nearly 50%. Water stress in the previous year significantly increases the probability of migration. Although our models explain only 5 to 11% of the variation in female migration, these results highlight the importance of structural determinants (education, household composition), and environmental factors (climatic conditions) in migration decisions, independent of the intervention under study. 



```{r}
#| label: fig-migrant-2x2
#| fig-cap: "2×2 DiD estimates for migration outcomes"

# Extract DID coefficients and confidence intervals

did_row <- function(model, year_post, term = "treat_post") {
  tidy(model) |>
    filter(term == !!term) |>
    transmute(
      year = year_post,
      estimate,
      se = std.error,
      ci_lower = estimate - 1.96 * std.error,
      ci_upper = estimate + 1.96 * std.error
    )
}

# Combine placebo and main effects
migrant_2x2_df <- bind_rows(
  did_row(migrant_pre, 2008),      # Placebo: 1997-2008
  did_row(migrant_main, 2021)      # Treatment: 2008-2021
) |>
  mutate(
    period = factor(
      c("Placebo (1997–2008)", "Treatment (2008–2021)"),
      levels = c("Placebo (1997–2008)", "Treatment (2008–2021)")
    )
  )

# Visualization

ggplot(migrant_2x2_df, aes(x = period, y = estimate, color = period)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", size = 0.8) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.15, size = 1) +
  geom_point(size = 4) +
  scale_color_manual(
    values = c("Placebo (1997–2008)" = "#E69F00", "Treatment (2008–2021)" = "#56B4E9"),
    guide = "none"
  ) +
  labs(
    x = NULL,
    y = "Difference-in-Differences Effect",
    title = "Impact on Migration: Placebo and Treatment Estimates",
    subtitle = "Effect sizes with 95% confidence intervals"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11, color = "gray30"),
    axis.text.x = element_text(size = 12)
  )
```


```{r}
# Women analysis 

pre_migrant <- dat_migrant |>
  filter(DHSYEAR %in% c(1997, 2008)) |>
  mutate(
    post = as.integer(DHSYEAR == 2008),
    treat_post = treat * post
  )

 migrant_pre_wm <- feols(
  migrant ~ treat + post + treat_post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = pre_migrant |> filter(sexe == "femme"),
  weights = ~w_all,
  cluster = ~cluster_uid
)

# Main treatment effect: Women only
main_migrant_wm <- feols(
  migrant ~ treat + post + treat_post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = main_migrant |> filter(sexe == "femme"),
  weights = ~w_all,
  cluster = ~cluster_uid
)

etable(
  migrant_pre_wm,
  main_migrant_wm,
  headers = c(
    "Placebo 1997-2008\n(Women only)",
    "Treatment 2008-2021\n(Women only)"
  ),
  title = "DID Estimates: Migration Effects (Women only)"
)
```

Our DID analysis specifically for women reveals that there is no statistically significant effect of treatment on the probability of migration. The placebo test for the period 1997-2008 yields a small, non significant coefficient (60.0285) and an estimated positive effect for the period 2008-2021 (+0.0759). The results also show that higher education remains the major determinant, increasing the probability of migration by 75% in the placebo group and 50% in the treatment group, while water stress in the previous year and household structure (male-headed households) significantly reduce the probability of migration. However, when the analysis is restricted to women, the statistical power is also reduced. 


```{r}
#| label: fig-migrant-2x2 women
#| fig-cap: "2×2 DiD estimates for migration outcomes (women only)"


# Extract DID coefficients and confidence intervals (Women)

did_row <- function(model, year_post, term = "treat_post") {
  tidy(model) |>
    filter(term == !!term) |>
    transmute(
      year = year_post,
      estimate,
      se = std.error,
      ci_lower = estimate - 1.96 * std.error,
      ci_upper = estimate + 1.96 * std.error
    )
}

# Combine placebo and main effects (Women only)
migrant_2x2_women_df <- bind_rows(
  did_row(migrant_pre_wm, 2008),    # Placebo: 1997-2008 (Women)
  did_row(main_migrant_wm, 2021)    # Treatment: 2008-2021 (Women)
) |>
  mutate(
    period = factor(
      c("Placebo (1997–2008)", "Treatment (2008–2021)"),
      levels = c("Placebo (1997–2008)", "Treatment (2008–2021)")
    )
  )


# Visualization

ggplot(migrant_2x2_women_df, aes(x = period, y = estimate, color = period)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", size = 0.8) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.15, size = 1) +
  geom_point(size = 4) +
  scale_color_manual(
    values = c("Placebo (1997–2008)" = "#E69F00", "Treatment (2008–2021)" = "#56B4E9"),
    guide = "none"
  ) +
  labs(
    x = NULL,
    y = "Difference-in-Differences Effect",
    title = "Impact on Migration: Placebo and Treatment Estimates (Women)",
    subtitle = "Effect sizes with 95% confidence intervals"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11, color = "gray30"),
    axis.text.x = element_text(size = 12)
  )

```



### Staggered DiD — static estimation

We use the Gardner (2022) two-stage DiD estimator (`did2s`) to leverage all available years with migration data (1997, 2008, 2018, 2021). This estimator is robust to heterogeneous treatment effects and staggered adoption.

```{r}
#| label: migrant-staggered-static

# Prepare relative year bins (same scheme as main analysis)
dat_migrant_did2s <- dat_migrant |>
  mutate(
    rel_year = DHSYEAR - STATUS_YR,
    rel_year_bin = case_when(
      is.infinite(rel_year)        ~ "never",
      is.na(rel_year)               ~ "never",
      rel_year <= -10                ~ "pre_lointain",
      rel_year >= -9 & rel_year <= -1 ~ "pre_proche",
      rel_year >=  0 & rel_year <=  9 ~ "post_proche",
      rel_year >= 10                ~ "post_lointain",
      TRUE                           ~ NA_character_
    ),
    rel_year_bin = factor(
      rel_year_bin,
      levels = c("pre_lointain", "pre_proche", "post_proche", "post_lointain", "never")
    ),
    treat_on = as.integer(treat == 1 & DHSYEAR >= STATUS_YR)
  )


did2s_static_migration <- did2s(
  data         = dat_migrant_did2s,
  yname        = "migrant",
  first_stage  = ~ sexe + education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(treat_on, ref = FALSE),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)
 
modelsummary(list("did2s migration (static, pooled men + women, 4 vagues)" = did2s_static_migration))

```

The static DID estimation suggests a positive, but statistically insignificant, association between treatment and the probability of migration (+3.5 percentage points, standard error of 2.2). Given the magnitude of the uncertainty, this estimates does not allow us to reject the null hypothesis of no effect, which justifies the use of a dynamic event study analysis. 



### Staggered DiD — event study

```{r}
#| label: migration-event-study

#| label: migration-event-study

did2s_migration_es <- did2s(
  data         = dat_migrant_did2s,
  yname        = "migrant",
  first_stage  = ~ sexe + education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_bin, ref = c("pre_proche", "never")),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

etable(did2s_migration_es, headers = "did2s migration event study")


```

Unlike the static DID analysis, the event study reveals temporal heterogeneity in the treatment effect on migration (for men and women combined). Long before treatment, no significant difference is found prior to the intervention. However, immediately after treatment, the event study reveals a positive and significant effect (+5.2), but in the long term, the effect disappears. The coefficient approaches zero and becomes insignificant, suggesting that the treated and control regions converge toward the same level of migration. This positive effect can be explained by the fact the protected area creation may have disrupted local economic activities, temporarily prompting people to seek opportunities elsewhere. Conversely, it may have improved local economic conditions just enough to allow people to migrate, or it may have simply provided a period for the population to adapt before turning to normal. The graph below illustrates this variation in effects. 


```{r}
bin_label <- function(k) {
  dplyr::case_when(
    is.na(k)             ~ NA_character_,
    k == "pre_lointain"  ~ "Pre (>1 vague d'ecart)",
    k == "post_proche"   ~ "Post (meme intervalle)",
    k == "post_lointain" ~ "Post (>1 vague d'ecart)",
    TRUE                 ~ as.character(k)
  )
}
 
event_study_df <- tidy(did2s_migration_es, conf.int = TRUE) |>
  filter(grepl("^rel_year_bin::", term)) |>
  mutate(
    k = sub("^rel_year_bin::", "", term),
    k = factor(k, levels = c("pre_lointain", "post_proche", "post_lointain"))
  ) |>
  arrange(k)

 
ggplot(event_study_df, aes(k, estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  labs(x = NULL,
       y = "Effect on the probability of migration",
       title = "Event study: migration around the creation of protected areas") +
  scale_x_discrete(labels = bin_label) +
  theme_minimal(base_size = 13)



```


## Migration by PA type - 2×2 DiD

We test whether effects differ between strict PAs (IUCN I–IV) and multi-use PAs (IUCN V–VI), following the same approach as the main analysis (H3). Given that migration is only observable on 4 waves (2008, 2018, 2021, only 1997 is women-only), we use the pooled sample (men + women, with a sex as a covariate) as the main specification here, consistent with the main migration DID. The women-only panel (`dat_migrant_femmes`)/ (`dat_migrant2_femmes`) is reported alongside as a robustness check, not as the primary estimate. 

```{r}
#| label: migrant-iucn-2x2

# Définir les fonctions
is_strict <- function(iucn_cat) {
  iucn_cat %in% c("Ia", "Ib", "II")  # Strictes
}

is_multi <- function(iucn_cat) {
  iucn_cat %in% c("IV", "V", "VI")   # Multi-usages
}


dat_migrant_iucn <- dat_migrant |>
  mutate(
    treat_strict = as.integer(GROUP == "Treatment" & is_strict(IUCN_CAT)),
    treat_multi  = as.integer(GROUP == "Treatment" & is_multi(IUCN_CAT))
  )


# Formule commune
formula_did <- ~ treat + post + treat:post + education +
                  spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220

# PLACEBO (1997-2008) 

# Strict placebo
pre_strict_migrant <- dat_migrant_iucn |>
  filter(DHSYEAR %in% c(1997, 2008), treat_strict == 1 | GROUP == "Control") |>
  mutate(post = as.integer(DHSYEAR == 2008))

iucn_strict_pre <- feols(
  migrant ~ treat_strict + post + treat_strict:post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = pre_strict_migrant, weights = ~w_all, cluster = ~cluster_uid
)

# Multi placebo
pre_multi_migrant <- dat_migrant_iucn |>
  filter(DHSYEAR %in% c(1997, 2008), treat_multi == 1 | GROUP == "Control") |>
  mutate(post = as.integer(DHSYEAR == 2008))

iucn_multi_pre <- feols(
  migrant ~ treat_multi + post + treat_multi:post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = pre_multi_migrant, weights = ~w_all, cluster = ~cluster_uid
)

# TREATMENT (2008-2021)

# Strict treatment
main_strict_migrant <- dat_migrant_iucn |>
  filter(DHSYEAR %in% c(2008, 2021), treat_strict == 1 | GROUP == "Control") |>
  mutate(post = as.integer(DHSYEAR == 2021))

iucn_strict_main <- feols(
  migrant ~ treat_strict + post + treat_strict:post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = main_strict_migrant, weights = ~w_all, cluster = ~cluster_uid
)

# Multi treatment
main_multi_migrant <- dat_migrant_iucn |>
  filter(DHSYEAR %in% c(2008, 2021), treat_multi == 1 | GROUP == "Control") |>
  mutate(post = as.integer(DHSYEAR == 2021))

iucn_multi_main <- feols(
  migrant ~ treat_multi + post + treat_multi:post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = main_multi_migrant, weights = ~w_all, cluster = ~cluster_uid
)

etable(iucn_strict_pre, iucn_strict_main, iucn_multi_pre, iucn_multi_main,
       headers = c("Strict plac.", "Strict treat.", "Multi plac.", "Multi treat."))
```


No category of protected area has a significant effect on migration. The treatment x post interaction coefficients are close to zero. This can be interpreted to mean that protected areas neither reduce migration nor stimulate it (no positive or negative effect was detected). Instead, the major determinants of migration and household needs, rather than the presence or absence of protected areas. 


```{r}
#| label: fig-migrant-iucn-2x2
#| fig-cap: "2×2 DiD for migration by PA type"

# Définir le positionnement pour éviter la superposition
pd <- position_dodge(width = 0.2)

migrant_iucn_df <- bind_rows(
  did_row(iucn_strict_pre, 2008, "treat_strict:post") |> mutate(type = "Strict"),
  did_row(iucn_strict_main, 2021, "treat_strict:post") |> mutate(type = "Strict"),
  did_row(iucn_multi_pre, 2008, "treat_multi:post") |> mutate(type = "Multi-use"),
  did_row(iucn_multi_main, 2021, "treat_multi:post") |> mutate(type = "Multi-use")
) |>
  mutate(
    period = factor(ifelse(year == 2008, "1997–2008", "2008–2021"),
                    levels = c("1997–2008", "2008–2021")),
    lo = estimate - 1.96 * se,
    hi = estimate + 1.96 * se
  )

ggplot(migrant_iucn_df, aes(x = period, y = estimate, colour = type)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 3, position = pd) +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.2, position = pd) +
  labs(x = NULL, y = "DiD effect on Pr(migration)",
       colour = "PA type",
       title = "Migration: 2×2 DiD by PA type") +
  theme_minimal(base_size = 13)
```


### Migration by PA type — staggered static

```{r}
#| label: migrant-iucn-staggered

# Strict
pre_strict_migrant <- dat_migrant_iucn |>
  filter(GROUP == "Control" | (GROUP == "Treatment" & is_strict(IUCN_CAT))) |>
  mutate(treat_on = as.integer(treat == 1 & DHSYEAR >= STATUS_YR))

iucn_strict_staggered <- did2s(
  data         = pre_strict_migrant,
  yname        = "migrant",
  first_stage  = ~ education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(treat_on, ref = FALSE),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

# Multi-usage
pre_multi_migrant <- dat_migrant_iucn |>
  filter(GROUP == "Control" | (GROUP == "Treatment" & is_multi(IUCN_CAT))) |>
  mutate(treat_on = as.integer(treat == 1 & DHSYEAR >= STATUS_YR))

iucn_multi_staggered <- did2s(
  data         = pre_multi_migrant,
  yname        = "migrant",
  first_stage  = ~ education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(treat_on, ref = FALSE),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

modelsummary(list(
  "Strict (Staggered)" = iucn_strict_staggered,
  "Multi (Staggered)"  = iucn_multi_staggered
))

```


### Migration by PA type — event study

```{r}
#| label: migrant-iucn-event-study

# Strict
pre_strict_migrant_es <- dat_migrant_iucn |>
  filter(GROUP == "Control" | (GROUP == "Treatment" & is_strict(IUCN_CAT))) |>
  mutate(
    treat_on = as.integer(treat == 1 & DHSYEAR >= STATUS_YR),
    rel_year_2yr2 = as.factor(DHSYEAR - STATUS_YR)
  )

# Vérifier les valeurs
print(table(pre_strict_migrant_es$rel_year_2yr2, useNA = "ifany"))

# Supprimer NA
pre_strict_migrant_es <- pre_strict_migrant_es |>
  filter(!is.na(rel_year_2yr2))

iucn_strict_es <- did2s(
  data         = pre_strict_migrant_es,
  yname        = "migrant",
  first_stage  = ~ education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_2yr2, ref = "-1"), 
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

```



```{r}
# covariate on the full sample, which is not the right comparator here).
did2s_migrant_es <- did2s(
  data         = dat_migrant_iucn,
  yname        = "migrant",
  first_stage  = ~ education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_2yr2, ref = c("-1", "never")),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

es_migrant_iucn <- bind_rows(
  tidy_es(did2s_migrant_es_strict, "Strict"),
  tidy_es(did2s_migrant_es_multi,  "Multi-use"),
  tidy_es(did2s_migrant_es_femmes, "Both")
)

ggplot(es_migrant_iucn, aes(k, estimate, colour = type)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(position = pd2, size = 2) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, position = pd2) +
  labs(x = "Years to treatment (2-year bins)",
       y = "Effect on Pr(migration)",
       colour = "PA type",
       title = "Event study: migration by PA type") +
  scale_x_discrete(labels = bin_label) +
  theme_minimal(base_size = 13)
```


## Summary of results

```{r}
summary_table <- bind_rows(
  extract_did2s(did2s_migrant_static_femmes, "Migration — All PAs (women)"),
  extract_did2s(did2s_migrant_strict, "Migration — Strict (women)"),
  extract_did2s(did2s_migrant_multi, "Migration — Multi-use (women)")
)
```


