---
output: typst
editor_options: 
  chunk_output_type: console
---

# Migration outcomes: household mobility

## Motivation

In the pre-analysis plan, we chose not to include migration among the control variables, as it is likely influenced both by the direct effects of protected areas and by broader changes in living standards in surrounding areas, making it a collider rather than a valid control. Instead, migration is treated as an outcome variable, since it is more plausibly a consequence of protected areas than an independent cause. This allows us to assess whether populations leave their territory because of protected areas, or whether migration is one of the mechanisms through which these areas affect living standard, without introducing bias into the analysis. In their paper, Jones et al. [-@Jones2018] state that human migration is often considered an important driver of land use change and a threat to protected area integrity. Understanding household mobility and behaviour at the forest frontier is important for guiding conservation and development policies.

## Variable definition

Among the three available data sources (DHS, MIS, MICS), the indicators that may help explain the movements of household members are as follows:

| Indicateur | Source | Level | Indicator definition |
|------------------|------------------|------------------|------------------|
| V104 | DHS/MIS (women) | Individual | Number of years the respondent has lived in the village, town, or city where she was interviewed. Visitors to the community are coded 96 |
| V135 / MV135 | DHS/MIS | Individual | Whether the respondent is a usual resident or just a visitor to the household |
| V136 / MV136 | DHS/MIS | Household | Total number of household members (residents + visitors who slept there) |
| MV104 | DHS/MIS (men) | Individual | Number of years the respondent has lived in the village, town, or city where he was interviewed. Visitors to the community are coded 96 |
| WB15 | MICS (women) | Individual | MICS equivalent of V104 |
| MWB15 | MICS (men) | Individual | MICS equivalent of V104 |
| HV102 | DHS/MIS | Household | De jure resident status (usual resident) |
| HV103 | DHS/MIS | Household | De facto resident status (slept there the previous night) |


## Data availability summary

We use the following indicators

| Year | Source   | V104 | MV104 | WB15 | MWB15 |
|------|----------|------|-------|-------|------|
| 1997 | DHS-III  | ✓    | ✗     | ✗     | ✗    |
| 2008 | DHS-V    | ✓    | ✓     | ✗     | ✗    |
| 2011 | MIS      | ✓    | ✗     | ✗     | ✗    |
| 2013 | MIS      | ✓    | ✗     | ✗     | ✗    |
| 2016 | MIS      | ✓    | ✗     | ✗     | ✗    |
| 2018 | MICS-6   | ✗    | ✗     | ✓     | ✓    |
| 2021 | DHS-VIII | ✓    | ✓     | ✗     | ✗    |


## Data loading and harmonization

```{r}

#| label: setup
#| message: false
#| warning: false


library(tidyverse)
library(haven)
library(purrr)
library(fixest)
library(survey)
library(did2s)
library(broom)
library(ggplot2)
library(gt)
library(sf)
library(modelsummary)
library(labelled)


options(scipen = 999)

options(survey.lonely.psu = "adjust")
```

## Load matched household data

```{r}
#| label: load-matched

load_matched <- function(path, year, spei_years) {
  read_rds(path) |>
    st_drop_geometry() |>
    mutate(DHSYEAR = year) |>
    rename(
      spei_wc_n_2 = !!sym(paste0("spei_wc_", spei_years[1])),
      spei_wc_n_1 = !!sym(paste0("spei_wc_", spei_years[2])),
      spei_wc_n   = !!sym(paste0("spei_wc_", spei_years[3]))
    ) |>
    mutate(
      hv219 = zap_labels(hv219),
      hv220 = as.numeric(zap_labels(hv220))
    ) |>
    filter(GROUP %in% c("Treatment", "Control")) |>
    select(DHSYEAR, hv001, hv002, hv005, GROUP, WDPAID, STATUS_YR, IUCN_CAT,
           treatment, weights, spei_wc_n_2, spei_wc_n_1, spei_wc_n, hv219, hv220)
}

matched_all <- bind_rows(
  load_matched("data/derived/data_matched_1997.rds", 1997, c(1995, 1996, 1997)),
  load_matched("data/derived/data_matched_2008.rds", 2008, c(2006, 2007, 2008)),
  load_matched("data/derived/data_matched_2011.rds", 2011, c(2009, 2010, 2011)),
  load_matched("data/derived/data_matched_2013.rds", 2013, c(2011, 2012, 2013)),
  load_matched("data/derived/data_matched_2016.rds", 2016, c(2014, 2015, 2016)),
  load_matched("data/derived/data_matched_2018.rds", 2018, c(2016, 2017, 2018)),
  load_matched("data/derived/data_matched_2021.rds", 2021, c(2019, 2020, 2021))
)

# Vérification avant jointure 

matched_all |> count(DHSYEAR, GROUP)
matched_all |> summarise(prop_status_yr_na = mean(is.na(STATUS_YR)))
```


## Load migration's survey data

### Définition du statut de résidence et de la migration 

```{r}
# Nettoyage de la durée de résidence -----------------------------------
# DHS/MIS (v104/mv104, 1997/2008/2011/2013/2016/2021)
# 95 = toujours vécu ici | 96 = visiteur | 97 = Inconsistent | 98 = Don't know | 99 = ne sait pas
# MICS (wb15/mwb15, 2018)
# 99 = Non reponse 

clean_duration <- function(x) {
  y <- as.numeric(zap_labels(x))
  y <- ifelse(y %in% c(97, 98, 99), NA_real_, y)
  y
}

# Identification du statut de résidence 

derive_migrant <- function(duree){
  
  statut_residence <- case_when(
    duree == 95 ~ "toujours ici",
    duree == 96 ~ "visiteur",
    duree >= 0 & duree < 95 ~ "duree_connue",
    TRUE ~ NA_character_
  )
  
  # Indicateur de migration 
  migrant <- case_when(
    duree == 95 ~ 0L,              # non migrant 
    duree == 96 ~ NA_integer_,     # visiteur
    duree >= 0 & duree < 95 ~ 1L,  # migrant
    TRUE ~ NA_integer_
  )
  
  # Indicateur de migration récente (migration depuis 5 ans ou moins)
  migrant_recent <- case_when(
    duree == 96 ~ NA_integer_,     # Visiteur
    duree == 95 ~ NA_integer_,     # non-migrant
    duree >= 0 & duree <= 5 ~ 1L,  # migrant récent
    duree > 5 & duree < 95 ~ 0L,    # migrant ancien
    TRUE ~ NA_integer_
  )
  
  tibble(
    statut_residence = statut_residence, 
    migrant = migrant, 
    migrant_recent = migrant_recent
    )
}

# Identification du type de migration 
derive_type_migration <- function(
    migrant,
    residence_precedente) {
  
  case_when(
    migrant == 1 & residence_precedente == "urbain"   ~ "urbain_vers_rural",
    migrant == 1 & residence_precedente == "rural"    ~ "rural_vers_rural",
    migrant == 1 & residence_precedente == "etranger" ~ "etranger_vers_rural",
    migrant == 0 ~ "non_migrant",
    TRUE ~ NA_character_
  )
}
```


### Harmonization of the education variable (v106 vs WB6A/MWB6A)

The variables related to education in the three data sources are not categorized in the same way. In the DHS/MIS Data, variable v106 uses only 4 categories (0 = No education, 1 = Primary, 2 = Secondary, 3 = Higher), whereas in the MICS data, variable WB6A for women and MWB6A for men use more categories (0 = Preschool, 1 = Primary, 2 = Seconday 1, 3 = Secondary 2, 4 = Higher, 8 = Don't know, 9 = No response). We will therefore harmonize these variables into 5 categories (None, Preschool, Primary, Secondary, Higher). 

```{r}
# Fonction 

education_labels <- c("Aucune", "Prescolaire", "Primaire", "Secondaire", "Supérieur")

recode_education_dhs <- function(v106) {
  x <- as.numeric(zap_labels(v106))
  case_when(
    x == 0 ~ 0L,  # Aucune
    x == 1 ~ 2L,  # Primaire
    x == 2 ~ 3L,  # Secondaire
    x == 3 ~ 4L,  # Superieur
    TRUE   ~ NA_integer_
  )
}

recode_education_mics <- function(wbx){
  x <- as.numeric(zap_labels(wbx))
  case_when(
    x == 0        ~ 1L,          # prescolaire 
    x == 1        ~ 2L,          # primaire
    x %in% c(2,3) ~ 3L,          # secondaire 1 + 2 -> secondaire
    x == 4        ~ 4L,          # superieur
    x %in% c(8,9) ~ NA_integer_, # NSP / pas de reponse
    TRUE          ~ NA_integer_
  )
}

var_education <- function(code){
  factor(code, levels = 0:4, labels = education_labels)
}
```


### Harmonization of place of residence (v105/mv105 vs WB16A/MWB16A)

The variables related to respondent's place of residence in the three data sources are not categorized in the same way. In the DHS/MIS Data, variable v105/mv105 uses these categories (0 = capital, large city, 1 = Small city, 2 = Town, 3 = Countryside, 4 = Abroad), whereas in the MICS data, variable WB16A for women and MWB16A for men use more categories ( 1 = Large city, 2 = Other city, 3 = rural area, 9 = No response). We will therefore harmonize these variables into 5 categories (None, Preschool, Primary, Secondary, Higher). 

#### Women Data

```{r}

# DHS 

lieu_precedent_dhs <- function(v105){
  x <- as.numeric(zap_labels(v105))
  case_when(
       x %in% c(0, 1, 2) ~ "urbain",   # Capital/large city, Small city, Town
    x == 3            ~ "rural",    # Countryside
    x == 4            ~ "etranger", # Abroad
    TRUE              ~ NA_character_
  )
}

# MICS

lieu_precedent_mics <- function(WB16A){
    x <- as.numeric(zap_labels(WB16A))
    case_when(
      x %in% c(1, 2) ~ "urbain", # Grande ville, Autre ville
      x == 3         ~ "rural",  # Milieu rural
      x == 9         ~ NA_character_, # Non reponse
      TRUE           ~ NA_character_
    )
}

```


### DHS/MIS files (1997, 2008, 2011, 2013, 2016, 2021)

The sampling stratum variable is designated as "v022" for the year 1997, and "v023" for all other DHS/MIS waves (2008, 2011, 2013, 2016, 2021). We rename the "sample stratum" variable to v023.  

#### Women

```{r}

# Helper DHS/MIS, Femmes-----------------------------
# v001: cluster number
# v002: household number 
# v005: sample weight
# v012: respondent's age
# v021: Primary sample unit
# v023: Sample strate
# v024: De facto region of residence 
# v025: De facto type of place of residence
# v104: Number of years the respondent has lived
# v105: Type of place of previous residence 
# v106: Highest education level attended


load_migration_ir <- function(path, year, strate_var_source,
                              extra_vars = c("v012", "v106"),
                              filter_rural = TRUE) {
  
  cols <- c(
    "v001", "v002", "v005", "v021", strate_var_source,
    "v024", "v025", "v026", "v104", "v105", extra_vars
  )
  
  df <- read_dta(
    path,
    col_select = all_of(unique(cols))
  )
  
  if (strate_var_source != "v023") {
    df <- df |> rename(v023 = !!sym(strate_var_source))
  }
  
  if (filter_rural) {
    df <- df |> filter(v025 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "femme",
      duree_residence = clean_duration(v104),
      
      residence_precedente = case_when(
        v104 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_dhs(v105)
      ),
      
      region_precedente = NA_real_,
      
      taille_localite = as.numeric(zap_labels(v026)),
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      education = var_education(
        recode_education_dhs(v106)
      )
    )
  
  df |>
    select(
      v001, v002, v005, v021, v023,
      DHSYEAR, sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025, v024,
      all_of(extra_vars)
    )
}

ir_1997 <- load_migration_ir("data/raw/dhs/DHS_1997/MDIR31DT/MDIR31FL.DTA", 1997, "v023")
ir_2008 <- load_migration_ir("data/raw/dhs/DHS_2008/MDIR51DT/MDIR51FL.DTA", 2008, "v023")
ir_2011 <- load_migration_ir("data/raw/dhs/DHS_2011/MDIR61DT/MDIR61FL.DTA", 2011, "v023")
ir_2013 <- load_migration_ir("data/raw/dhs/DHS_2013/MDIR6ADT/MDIR6AFL.DTA", 2013, "v023")
ir_2016 <- load_migration_ir("data/raw/dhs/DHS_2016/MDIR71DT/MDIR71FL.DTA", 2016, "v023")
ir_2021 <- load_migration_ir("data/raw/dhs/DHS_2021/MDIR81DT/MDIR81FL.DTA", 2021, "v023")

```

#### Men Data

```{r}
# Helper DHS/MIS, hommes ----------------------------------------------
# mv012: respondent's age
# mv106: Highest education level attended
# mv024: De facto region of residence 
# mv025: De facto type of place of residence
# mv001: cluster number
# mv002: household number 
# mv005: sample weight
# mv025: Primary sample unit
# mv023: Sample strate
# mv104: Number of years the respondent has lived
# mv105: Type of place of previous residence 

load_migration_mr <- function(path, year, filtrer_rural = TRUE) {
  
  cols <- c(
    "mv001", "mv002", "mv005", "mv021", "mv023",
    "mv104", "mv105", "mv025", "mv026",
    "mv012", "mv106", "mv024"
  )
  
  df <- read_dta(
    path,
    col_select = all_of(cols)
  ) |>
    rename(
      v001 = mv001,
      v002 = mv002,
      v005 = mv005,
      v021 = mv021,
      v023 = mv023,
      v012 = mv012,
      v106 = mv106,
      v024 = mv024,
      v025 = mv025
    )
  
  if (filtrer_rural) {
    df <- df |> filter(v025 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "homme",
      
      duree_residence = clean_duration(mv104),
      
      residence_precedente = case_when(
        mv104 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_dhs(mv105)
      ),
      
      region_precedente = NA_real_,
      
      taille_localite = as.numeric(zap_labels(mv026)),
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      education = var_education(
        recode_education_dhs(v106)
      )
    )
  
  df |>
    select(
      v001, v002, v005, v021, v023,
      DHSYEAR, sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025, v012, v106, v024
    )
}

#Homme
#mr_1997 <- load_migration_mr("data/raw/dhs/DHS_1997/MDPR31DT/MDPR31FL.DTA", 1997, "DHS-III")
mr_2008 <- load_migration_mr("data/raw/dhs/DHS_2008/MDMR51DT/MDMR51FL.DTA", 2008)
#mr_2011 <- load_migration_mr("data/raw/dhs/DHS_2011/MDIR61DT/MDIR61FL.DTA", 2011, "MIS")
#mr_2013 <- load_migration_mr("data/raw/dhs/DHS_2013/MDMR6ADT/MDMR6AFL.DTA", 2013, "MIS")
#mr_2016 <- load_migration_mr("data/raw/dhs/DHS_2016/MDMR71DT/MDMR71FL.DTA", 2016, "MIS")
mr_2021 <- load_migration_mr("data/raw/dhs/DHS_2021/MDMR81DT/MDMR81FL.DTA", 2021)

```


### MICS Data (2018)

#### Women data 

```{r}

# Femme--------------------------------------------------

# v001/mv001: cluster number
# v002/mv002: household number 
# v005/mv005: sample weight
# v012/mv012: respondent's age
# v021/mv025: Primary sample unit
# v023/mv023: Sample strate
# v024/mv024: De facto region of residence 
# v025/mv025: De facto type of place of residence
# v104/mv104: Number of years the respondent has lived
# v105/mv105: Type of place of previous residence 
# v106/mv106: Highest education level attended
# mnweight: sample weight

load_migration_mics_femme <- function(path, year = 2018, filter_rural = TRUE) {
  
  df <- read_sav(
    path,
    col_select = c(
      HH1, HH2, stratum, wmweight,
      WB4, WB6A, WB15, WB16A, WB17,
      welevel, HH6, HH7
    )
  )
  
  if (filter_rural) {
    df <- df |> filter(HH6 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "femme",
      
      duree_residence = clean_duration(WB15),
      
      residence_precedente = case_when(
        WB15 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_mics(WB16A)
      ),
      
      region_precedente = as.numeric(zap_labels(WB17)),
      
      taille_localite = NA_real_,
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      
      education = var_education(
        recode_education_mics(WB6A)
      )
    )
  
  df |>
    transmute(
      v001 = HH1,
      v002 = as.numeric(HH2),
      v005 = wmweight,
      v021 = HH1,
      v023 = as.numeric(zap_labels(stratum)),
      DHSYEAR,
      sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025 = HH6,
      v012 = WB4,
      v106 = welevel,
      v024 = HH7
    )
}


wm_2018 <- load_migration_mics_femme("data/raw/mics/2018/SPSS datasets/wm.sav")

```


#### Men data

```{r}

# Homme-----------------------------------------------------------

# HH1: cluster number
# HH2: Household number 
# MWB4: respondent's age 
# WB6: Highest education level attended (Primaire = 1, Secondaire 1 = 2, Secondaire 2 = 3, Supérieur = 4)
# mwelevel: education 
# MWB15: Number of years the respondent has lived
# MWB16A: Type of place of previous residence
# MWB17: Region of previous residence
# mnweight: sample weight


load_migration_mics_homme <- function(path, year = 2018,
                                      filter_rural = TRUE) {
  
  df <- read_sav(
    path,
    col_select = c(
      HH1, HH2, HH6, HH7, stratum, mnweight,
      MWB4, MWB15, MWB16A, MWB17, MWB6A, mwelevel
    )
  )
  
  if (filter_rural) {
    df <- df |> filter(HH6 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "homme",
      
      duree_residence = clean_duration(MWB15),
      
      residence_precedente = case_when(
        MWB15 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_mics(MWB16A)
      ),
      
      region_precedente = as.numeric(zap_labels(MWB17)),
      
      taille_localite = NA_real_,
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      
      education = var_education(
        recode_education_mics(MWB6A)
      )
    )
  
  df |>
    transmute(
      v001 = HH1,
      v002 = as.numeric(HH2),
      v005 = mnweight,
      v021 = HH1,
      v023 = as.numeric(zap_labels(stratum)),
      DHSYEAR,
      sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025 = HH6,
      v012 = MWB4,
      v106 = mwelevel,
      v024 = HH7
    )
}


mn_2018 <- load_migration_mics_homme("data/raw/mics/2018/SPSS datasets/mn.sav")

```


##### Combiner les données en une seule base 

```{r}
# DHS------------------------------
dhs_only <- bind_rows(ir_1997, ir_2008, ir_2021, mr_2008, mr_2021)
 
# Verification : 1997 ne doit contenir que des femmes
dhs_only |> count(DHSYEAR, sexe)

# MIS------------------------------
mis_only <- bind_rows(ir_2011, ir_2013, ir_2016)

# MICS-----------------------------
mics_only <- bind_rows(wm_2018, mn_2018)

# Combiner tous les données 
migration_raw <- bind_rows(dhs_only, mis_only, mics_only)
 
migration_raw |> count(DHSYEAR, sexe)
migration_raw |> count(DHSYEAR, v025)


```



### Descriptive statistics 

























