---
output: typst
editor_options: 
  chunk_output_type: console
---

# Migration outcomes: household mobility

## Motivation

In the pre-analysis plan, we chose not to include migration among the control variables, as it is likely influenced both by the direct effects of protected areas and by broader changes in living standards in surrounding areas, making it a collider rather than a valid control. Instead, migration is treated as an outcome variable, since it is more plausibly a consequence of protected areas than an independent cause. This allows us to assess whether populations leave their territory because of protected areas, or whether migration is one of the mechanisms through which these areas affect living standard, without introducing bias into the analysis. In their paper, Jones et al. [-@Jones2018] state that human migration is often considered an important driver of land use change and a threat to protected area integrity. Understanding household mobility and behaviour at the forest frontier is important for guiding conservation and development policies.

## Variable definition

Among the three available data sources (DHS, MIS, MICS), the indicators that may help explain the movements of household members are as follows:

| Indicateur | Source | Level | Indicator definition |
|------------------|------------------|------------------|------------------|
| V104 | DHS/MIS (women) | Individual | Number of years the respondent has lived in the village, town, or city where she was interviewed. Visitors to the community are coded 96 |
| V135 / MV135 | DHS/MIS | Individual | Whether the respondent is a usual resident or just a visitor to the household |
| V136 / MV136 | DHS/MIS | Household | Total number of household members (residents + visitors who slept there) |
| MV104 | DHS/MIS (men) | Individual | Number of years the respondent has lived in the village, town, or city where he was interviewed. Visitors to the community are coded 96 |
| WB15 | MICS (women) | Individual | MICS equivalent of V104 |
| MWB15 | MICS (men) | Individual | MICS equivalent of V104 |
| HV102 | DHS/MIS | Household | De jure resident status (usual resident) |
| HV103 | DHS/MIS | Household | De facto resident status (slept there the previous night) |


## Data availability summary

We use the following indicators

| Year | Source   | V104 | MV104 | WB15 | MWB15 |
|------|----------|------|-------|-------|------|
| 1997 | DHS-III  | ✓    | ✗     | ✗     | ✗    |
| 2008 | DHS-V    | ✓    | ✓     | ✗     | ✗    |
| 2011 | MIS      | ✓    | ✗     | ✗     | ✗    |
| 2013 | MIS      | ✓    | ✗     | ✗     | ✗    |
| 2016 | MIS      | ✓    | ✗     | ✗     | ✗    |
| 2018 | MICS-6   | ✗    | ✗     | ✓     | ✓    |
| 2021 | DHS-VIII | ✓    | ✓     | ✗     | ✗    |


## Data loading and harmonization

```{r}

#| label: setup
#| message: false
#| warning: false


library(tidyverse)
library(haven)
library(purrr)
library(fixest)
library(survey)
library(did2s)
library(broom)
library(ggplot2)
library(gt)
library(sf)
library(modelsummary)
library(labelled)


options(scipen = 999)

options(survey.lonely.psu = "adjust")
```

## Load matched household data

```{r}
#| label: load-matched

load_matched <- function(path, year, spei_years) {
  read_rds(path) |>
    st_drop_geometry() |>
    mutate(DHSYEAR = year) |>
    rename(
      spei_wc_n_2 = !!sym(paste0("spei_wc_", spei_years[1])),
      spei_wc_n_1 = !!sym(paste0("spei_wc_", spei_years[2])),
      spei_wc_n   = !!sym(paste0("spei_wc_", spei_years[3]))
    ) |>
    mutate(
      hv219 = zap_labels(hv219),
      hv220 = as.numeric(zap_labels(hv220))
    ) |>
    filter(GROUP %in% c("Treatment", "Control")) |>
    select(DHSYEAR, hv001, hv002, hv005, GROUP, WDPAID, STATUS_YR, IUCN_CAT,
           treatment, weights, spei_wc_n_2, spei_wc_n_1, spei_wc_n, hv219, hv220)
}

matched_all <- bind_rows(
  load_matched("data/derived/data_matched_1997.rds", 1997, c(1995, 1996, 1997)),
  load_matched("data/derived/data_matched_2008.rds", 2008, c(2006, 2007, 2008)),
  load_matched("data/derived/data_matched_2011.rds", 2011, c(2009, 2010, 2011)),
  load_matched("data/derived/data_matched_2013.rds", 2013, c(2011, 2012, 2013)),
  load_matched("data/derived/data_matched_2016.rds", 2016, c(2014, 2015, 2016)),
  load_matched("data/derived/data_matched_2018.rds", 2018, c(2016, 2017, 2018)),
  load_matched("data/derived/data_matched_2021.rds", 2021, c(2019, 2020, 2021))
)

# Vérification avant jointure 

matched_all |> count(DHSYEAR, GROUP)
matched_all |> summarise(prop_status_yr_na = mean(is.na(STATUS_YR)))
```


## Load migration's survey data

### Définition du statut de résidence et de la migration 

```{r}
# Nettoyage de la durée de résidence -----------------------------------
# DHS/MIS (v104/mv104, 1997/2008/2011/2013/2016/2021)
# 95 = toujours vécu ici | 96 = visiteur | 97 = Inconsistent | 98 = Don't know | 99 = ne sait pas
# MICS (wb15/mwb15, 2018)
# 99 = Non reponse 

clean_duration <- function(x) {
  y <- as.numeric(zap_labels(x))
  y <- ifelse(y %in% c(97, 98, 99), NA_real_, y)
  y
}

# Identification du statut de résidence 

derive_migrant <- function(duree){
  
  statut_residence <- case_when(
    duree == 95 ~ "toujours ici",
    duree == 96 ~ "visiteur",
    duree >= 0 & duree < 95 ~ "duree_connue",
    TRUE ~ NA_character_
  )
  
  # Indicateur de migration 
  migrant <- case_when(
    duree == 95 ~ 0L,              # non migrant 
    duree == 96 ~ NA_integer_,     # visiteur
    duree >= 0 & duree < 95 ~ 1L,  # migrant
    TRUE ~ NA_integer_
  )
  
  # Indicateur de migration récente (migration depuis 5 ans ou moins)
  migrant_recent <- case_when(
    duree == 96 ~ NA_integer_,     # Visiteur
    duree == 95 ~ NA_integer_,     # non-migrant
    duree >= 0 & duree <= 5 ~ 1L,  # migrant récent
    duree > 5 & duree < 95 ~ 0L,    # migrant ancien
    TRUE ~ NA_integer_
  )
  
  tibble(
    statut_residence = statut_residence, 
    migrant = migrant, 
    migrant_recent = migrant_recent
    )
}

# Identification du type de migration 
derive_type_migration <- function(
    migrant,
    residence_precedente) {
  
  case_when(
    migrant == 1 & residence_precedente == "urbain"   ~ "urbain_vers_rural",
    migrant == 1 & residence_precedente == "rural"    ~ "rural_vers_rural",
    migrant == 1 & residence_precedente == "etranger" ~ "etranger_vers_rural",
    migrant == 0 ~ "non_migrant",
    TRUE ~ NA_character_
  )
}
```


### Harmonization of the education variable (v106 vs WB6A/MWB6A)

The variables related to education in the three data sources are not categorized in the same way. In the DHS/MIS Data, variable v106 uses only 4 categories (0 = No education, 1 = Primary, 2 = Secondary, 3 = Higher), whereas in the MICS data, variable WB6A for women and MWB6A for men use more categories (0 = Preschool, 1 = Primary, 2 = Seconday 1, 3 = Secondary 2, 4 = Higher, 8 = Don't know, 9 = No response). We will therefore harmonize these variables into 5 categories (None, Preschool, Primary, Secondary, Higher). 

```{r}
# Fonction 

education_labels <- c("Aucune", "Prescolaire", "Primaire", "Secondaire", "Supérieur")

recode_education_dhs <- function(v106) {
  x <- as.numeric(zap_labels(v106))
  case_when(
    x == 0 ~ 0L,  # Aucune
    x == 1 ~ 2L,  # Primaire
    x == 2 ~ 3L,  # Secondaire
    x == 3 ~ 4L,  # Superieur
    TRUE   ~ NA_integer_
  )
}

recode_education_mics <- function(wbx){
  x <- as.numeric(zap_labels(wbx))
  case_when(
    x == 0        ~ 1L,          # prescolaire 
    x == 1        ~ 2L,          # primaire
    x %in% c(2,3) ~ 3L,          # secondaire 1 + 2 -> secondaire
    x == 4        ~ 4L,          # superieur
    x %in% c(8,9) ~ NA_integer_, # NSP / pas de reponse
    TRUE          ~ NA_integer_
  )
}

var_education <- function(code){
  factor(code, levels = 0:4, labels = education_labels)
}
```


### Harmonization of place of residence (v105/mv105 vs WB16A/MWB16A)

The variables related to respondent's place of residence in the three data sources are not categorized in the same way. In the DHS/MIS Data, variable v105/mv105 uses these categories (0 = capital, large city, 1 = Small city, 2 = Town, 3 = Countryside, 4 = Abroad), whereas in the MICS data, variable WB16A for women and MWB16A for men use more categories ( 1 = Large city, 2 = Other city, 3 = rural area, 9 = No response). We will therefore harmonize these variables into 5 categories (None, Preschool, Primary, Secondary, Higher). 

#### Women Data

```{r}

# DHS 

lieu_precedent_dhs <- function(v105){
  x <- as.numeric(zap_labels(v105))
  case_when(
       x %in% c(0, 1, 2) ~ "urbain",   # Capital/large city, Small city, Town
    x == 3            ~ "rural",    # Countryside
    x == 4            ~ "etranger", # Abroad
    TRUE              ~ NA_character_
  )
}

# MICS

lieu_precedent_mics <- function(WB16A){
    x <- as.numeric(zap_labels(WB16A))
    case_when(
      x %in% c(1, 2) ~ "urbain", # Grande ville, Autre ville
      x == 3         ~ "rural",  # Milieu rural
      x == 9         ~ NA_character_, # Non reponse
      TRUE           ~ NA_character_
    )
}

```


### DHS/MIS files (1997, 2008, 2011, 2013, 2016, 2021)

The sampling stratum variable is designated as "v022" for the year 1997, and "v023" for all other DHS/MIS waves (2008, 2011, 2013, 2016, 2021). We rename the "sample stratum" variable to v023.  

#### Women

```{r}

# Helper DHS/MIS, Femmes-----------------------------
# v001: cluster number
# v002: household number 
# v005: sample weight
# v012: respondent's age
# v021: Primary sample unit
# v023: Sample strate
# v024: De facto region of residence 
# v025: De facto type of place of residence
# v104: Number of years the respondent has lived
# v105: Type of place of previous residence 
# v106: Highest education level attended


load_migration_ir <- function(path, year, strate_var_source,
                              extra_vars = c("v012", "v106"),
                              filter_rural = TRUE) {
  
  cols <- c(
    "v001", "v002", "v005", "v021", strate_var_source,
    "v024", "v025", "v026", "v104", "v105", extra_vars
  )
  
  df <- read_dta(
    path,
    col_select = all_of(unique(cols))
  )
  
  if (strate_var_source != "v023") {
    df <- df |> rename(v023 = !!sym(strate_var_source))
  }
  
  if (filter_rural) {
    df <- df |> filter(v025 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "femme",
      duree_residence = clean_duration(v104),
      
      residence_precedente = case_when(
        v104 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_dhs(v105)
      ),
      
      region_precedente = NA_real_,
      
      taille_localite = as.numeric(zap_labels(v026)),
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      education = var_education(
        recode_education_dhs(v106)
      )
    )
  
  df |>
    select(
      v001, v002, v005, v021, v023,
      DHSYEAR, sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025, v024,
      all_of(extra_vars)
    )
}

ir_1997 <- load_migration_ir("data/raw/dhs/DHS_1997/MDIR31DT/MDIR31FL.DTA", 1997, "v023")
ir_2008 <- load_migration_ir("data/raw/dhs/DHS_2008/MDIR51DT/MDIR51FL.DTA", 2008, "v023")
ir_2011 <- load_migration_ir("data/raw/dhs/DHS_2011/MDIR61DT/MDIR61FL.DTA", 2011, "v023")
ir_2013 <- load_migration_ir("data/raw/dhs/DHS_2013/MDIR6ADT/MDIR6AFL.DTA", 2013, "v023")
ir_2016 <- load_migration_ir("data/raw/dhs/DHS_2016/MDIR71DT/MDIR71FL.DTA", 2016, "v023")
ir_2021 <- load_migration_ir("data/raw/dhs/DHS_2021/MDIR81DT/MDIR81FL.DTA", 2021, "v023")

```

#### Men Data

```{r}
# Helper DHS/MIS, hommes ----------------------------------------------
# mv012: respondent's age
# mv106: Highest education level attended
# mv024: De facto region of residence 
# mv025: De facto type of place of residence
# mv001: cluster number
# mv002: household number 
# mv005: sample weight
# mv025: Primary sample unit
# mv023: Sample strate
# mv104: Number of years the respondent has lived
# mv105: Type of place of previous residence 

load_migration_mr <- function(path, year, filtrer_rural = TRUE) {
  
  cols <- c(
    "mv001", "mv002", "mv005", "mv021", "mv023",
    "mv104", "mv105", "mv025", "mv026",
    "mv012", "mv106", "mv024"
  )
  
  df <- read_dta(
    path,
    col_select = all_of(cols)
  ) |>
    rename(
      v001 = mv001,
      v002 = mv002,
      v005 = mv005,
      v021 = mv021,
      v023 = mv023,
      v012 = mv012,
      v106 = mv106,
      v024 = mv024,
      v025 = mv025
    )
  
  if (filtrer_rural) {
    df <- df |> filter(v025 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "homme",
      
      duree_residence = clean_duration(mv104),
      
      residence_precedente = case_when(
        mv104 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_dhs(mv105)
      ),
      
      region_precedente = NA_real_,
      
      taille_localite = as.numeric(zap_labels(mv026)),
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      education = var_education(
        recode_education_dhs(v106)
      )
    )
  
  df |>
    select(
      v001, v002, v005, v021, v023,
      DHSYEAR, sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025, v012, v106, v024
    )
}

#Homme
#mr_1997 <- load_migration_mr("data/raw/dhs/DHS_1997/MDPR31DT/MDPR31FL.DTA", 1997, "DHS-III")
mr_2008 <- load_migration_mr("data/raw/dhs/DHS_2008/MDMR51DT/MDMR51FL.DTA", 2008)
#mr_2011 <- load_migration_mr("data/raw/dhs/DHS_2011/MDIR61DT/MDIR61FL.DTA", 2011, "MIS")
#mr_2013 <- load_migration_mr("data/raw/dhs/DHS_2013/MDMR6ADT/MDMR6AFL.DTA", 2013, "MIS")
#mr_2016 <- load_migration_mr("data/raw/dhs/DHS_2016/MDMR71DT/MDMR71FL.DTA", 2016, "MIS")
mr_2021 <- load_migration_mr("data/raw/dhs/DHS_2021/MDMR81DT/MDMR81FL.DTA", 2021)

```


### MICS Data (2018)

#### Women data 

```{r}

# Femme--------------------------------------------------

# v001/mv001: cluster number
# v002/mv002: household number 
# v005/mv005: sample weight
# v012/mv012: respondent's age
# v021/mv025: Primary sample unit
# v023/mv023: Sample strate
# v024/mv024: De facto region of residence 
# v025/mv025: De facto type of place of residence
# v104/mv104: Number of years the respondent has lived
# v105/mv105: Type of place of previous residence 
# v106/mv106: Highest education level attended
# mnweight: sample weight

load_migration_mics_femme <- function(path, year = 2018, filter_rural = TRUE) {
  
  df <- read_sav(
    path,
    col_select = c(
      HH1, HH2, stratum, wmweight,
      WB4, WB6A, WB15, WB16A, WB17,
      welevel, HH6, HH7
    )
  )
  
  if (filter_rural) {
    df <- df |> filter(HH6 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "femme",
      
      duree_residence = clean_duration(WB15),
      
      residence_precedente = case_when(
        WB15 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_mics(WB16A)
      ),
      
      region_precedente = as.numeric(zap_labels(WB17)),
      
      taille_localite = NA_real_,
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      
      education = var_education(
        recode_education_mics(WB6A)
      )
    )
  
  df |>
    transmute(
      v001 = HH1,
      v002 = as.numeric(HH2),
      v005 = wmweight,
      v021 = HH1,
      v023 = as.numeric(zap_labels(stratum)),
      DHSYEAR,
      sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025 = HH6,
      v012 = WB4,
      v106 = welevel,
      v024 = HH7
    )
}


wm_2018 <- load_migration_mics_femme("data/raw/mics/2018/SPSS datasets/wm.sav")

```


#### Men data

```{r}

# Homme-----------------------------------------------------------

# HH1: cluster number
# HH2: Household number 
# MWB4: respondent's age 
# WB6: Highest education level attended (Primaire = 1, Secondaire 1 = 2, Secondaire 2 = 3, Supérieur = 4)
# mwelevel: education 
# MWB15: Number of years the respondent has lived
# MWB16A: Type of place of previous residence
# MWB17: Region of previous residence
# mnweight: sample weight


load_migration_mics_homme <- function(path, year = 2018,
                                      filter_rural = TRUE) {
  
  df <- read_sav(
    path,
    col_select = c(
      HH1, HH2, HH6, HH7, stratum, mnweight,
      MWB4, MWB15, MWB16A, MWB17, MWB6A, mwelevel
    )
  )
  
  if (filter_rural) {
    df <- df |> filter(HH6 == 2)
  }
  
  df <- df |>
    mutate(
      DHSYEAR = as.integer(year),
      sexe = "homme",
      
      duree_residence = clean_duration(MWB15),
      
      residence_precedente = case_when(
        MWB15 %in% c(95, 96) ~ NA_character_,
        TRUE ~ lieu_precedent_mics(MWB16A)
      ),
      
      region_precedente = as.numeric(zap_labels(MWB17)),
      
      taille_localite = NA_real_,
      
      annee_arrivee = ifelse(
        !is.na(duree_residence) & duree_residence < 95,
        DHSYEAR - duree_residence,
        NA_real_
      )
    )
  
  df <- bind_cols(
    df,
    derive_migrant(df$duree_residence)
  )
  
  df <- df |>
    mutate(
      type_migration = derive_type_migration(
        migrant,
        residence_precedente
      ),
      
      education = var_education(
        recode_education_mics(MWB6A)
      )
    )
  
  df |>
    transmute(
      v001 = HH1,
      v002 = as.numeric(HH2),
      v005 = mnweight,
      v021 = HH1,
      v023 = as.numeric(zap_labels(stratum)),
      DHSYEAR,
      sexe,
      duree_residence,
      statut_residence,
      migrant,
      migrant_recent,
      residence_precedente,
      region_precedente,
      type_migration,
      annee_arrivee,
      education,
      taille_localite,
      v025 = HH6,
      v012 = MWB4,
      v106 = mwelevel,
      v024 = HH7
    )
}


mn_2018 <- load_migration_mics_homme("data/raw/mics/2018/SPSS datasets/mn.sav")

```


### Combine the data into a single dataset 

```{r}
# DHS------------------------------
dhs_only <- bind_rows(ir_1997, ir_2008, ir_2021, mr_2008, mr_2021)
 
# Verification : 1997 ne doit contenir que des femmes
dhs_only |> count(DHSYEAR, sexe)

# MIS------------------------------
mis_only <- bind_rows(ir_2011, ir_2013, ir_2016)

# MICS-----------------------------
mics_only <- bind_rows(wm_2018, mn_2018)

# Combiner tous les données 
migration_pooled <- bind_rows(dhs_only, mis_only, mics_only)
 
migration_pooled |> count(DHSYEAR, sexe)
migration_pooled |> count(DHSYEAR, v025)


```

### Join with matched data

```{r}
migration_pooled <- migration_pooled |>
  left_join(matched_all, by = c("DHSYEAR" = "DHSYEAR", "v001" = "hv001", "v002" = "hv002"))
 
# voir si on a une prop_non_matche anormalement elevee ou non (proche de 1).
migration_pooled <- migration_pooled |>
  mutate(poids_brut = ifelse(DHSYEAR == 2018, v005, v005 / 1000000)) |>
  group_by(DHSYEAR, sexe) |>
  mutate(poids_norm = poids_brut / mean(poids_brut, na.rm = TRUE)) |>
  ungroup() |>
  mutate(
    w_all       = poids_norm * weights,
    cluster_uid = paste(DHSYEAR, v001, sep = "_"),
    treat       = as.integer(GROUP == "Treatment"),
    sexe_chef_menage = as_factor(hv219),
    age_chef_menage  = hv220
  )

```


## Descriptive statistics 

```{r}
#| label: setup-svy-design 
# Primary sampling unit =v021, Sample strate = v023, weights = w_all

migration_svy_data <- migration_pooled |>
  filter(!is.na(w_all), !is.na(v021), !is.na(v023), !is.na(GROUP))


design_migration <- svydesign(
  ids     = ~v021,
  strata  = ~v023,
  weights = ~w_all,
  data    = migration_svy_data,
  nest    = TRUE
)

```


```{r}
#| label: descriptive

# Statistiques pondérées (design-based) par année et par groupe
migrant_desc_svy <- svyby(
  ~migrant + migrant_recent,
  ~DHSYEAR + GROUP,
  design_migration,
  svymean,
  na.rm = TRUE,
  vartype = "ci"
) |>
  as_tibble() |>
  arrange(DHSYEAR, GROUP)

migrant_desc_svy

# Effectifs bruts 
migration_svy_data |> 
  summarise(n = n(), .by = c(DHSYEAR, GROUP))


# Type de migration par année et groupe (pondéré)
type_migration_desc_svy <- svytable(
  ~DHSYEAR + GROUP + type_migration,
  design_migration
) |>
  as_tibble() |>
  mutate(pct = n / sum(n) * 100, .by = c(DHSYEAR, GROUP))

type_migration_desc_svy
```



```{r}
#| label: fig-migration-rate
#| fig-cap: "Share of migrants by survey year and treatment group (weighted)"

ggplot(migrant_desc_svy, aes(x = factor(DHSYEAR), y = migrant * 100, fill = GROUP)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(
    aes(ymin = ci_l.migrant * 100, ymax = ci_u.migrant * 100),
    position = position_dodge(width = 0.7), width = 0.2
  ) +
  labs(x = "Survey year", y = "Share of migrants (%)", fill = "Group",
       title = "Migration prevalence by year and treatment status (weighted)") +
  theme_minimal(base_size = 13)
```




```{r}
#| label: tbl-balance
#| tbl-cap: "Équilibre des covariables entre groupe traité et groupe contrôle"

datasummary_balance(
  ~GROUP,
  data = migration_svy_data |>
    select(GROUP, age_chef_menage, education, sexe_chef_menage,
           duree_residence, taille_localite),
  output = "gt"
)
```




```{r}
#| label: fig-migration-recent
#| fig-cap: "Share of recent migrants by survey year and treatment group (weighted)"

ggplot(migrant_desc_svy, aes(x = factor(DHSYEAR), y = migrant_recent * 100, fill = GROUP)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(
    aes(ymin = ci_l.migrant_recent * 100, ymax = ci_u.migrant_recent * 100),
    position = position_dodge(width = 0.7), width = 0.2
  ) +
  labs(x = "Survey year",
       y = "Share of recent migrants (%)",
       fill = "Group",
       title = "Recent migration prevalence by year and treatment status (weighted)") +
  theme_minimal(base_size = 13)
```



```{r}
#| label: fig-migration-type
#| fig-cap: "Distribution of migration types by survey year and treatment group (weighted)"

ggplot(type_migration_desc_svy, aes(x = factor(DHSYEAR), y = pct, fill = type_migration)) +
  geom_col(position = "stack", width = 0.6) +
  facet_wrap(~GROUP) +
  labs(x = "Survey year",
       y = "Share (%)",
       fill = "Type de migration",
       title = "Composition des types de migration par année et groupe (pondéré)") +
  theme_minimal(base_size = 13)
```



## Migration analysis

### 2×2 DiD

We follow the same 2×2 DiD design as the main analysis: a placebo test (1997–2008) and the treatment estimation (2008–2021). The model is:

$$Migrant_{ict} = \alpha + \beta_1 \text{Treat}_c + \beta_2 \text{Post}_t + \delta (\text{Treat}_c \times \text{Post}_t) + \mathbf{X}_{ct}'\gamma + \varepsilon_{ict}$$

where $i$ indexes individuals (women or men depending on the specification), $c$ clusters (PSU), $t$ survey rounds. $Migrant_{ict}$ is a binary variable (moved to current place of residence vs. has always lived there). Controls $\mathbf{X}$ include SPEI values, household head sex and age, education level and the individual's gender. Standard errors are clustered at the cluster-year level and all estimates are weighted by the combined survey weights $w_{ict}$. 


```{r}
#| label: migrant-2x2

# Placebo 1997–2008
pre_migrant <- migration_pooled |>
  filter(DHSYEAR %in% c(1997, 2008)) |>
  mutate(post = as.integer(DHSYEAR == 2008),
         treat_post = treat * post)

# Women analysis 
m_migrant_pre_A <- feols(
  migrant ~ treat + post + treat_post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = pre_migrant |> filter(sexe == "femme"),
  weights = ~w_all, cluster = ~cluster_uid
)

# Sex as a covariate (full sample)
m_migrant_pre_B <- feols(
  migrant ~ treat + post + treat_post + sexe + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = pre_migrant, weights = ~w_all, cluster = ~cluster_uid
)

# Treatment 2008–2021
main_migrant <- migration_pooled |>
  filter(DHSYEAR %in% c(2008, 2021)) |>
  mutate(post = as.integer(DHSYEAR == 2021),
         treat_post = treat * post)

m_migrant_main <- feols(
  migrant ~ treat + post + treat_post + sexe + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = main_migrant, weights = ~w_all, cluster = ~cluster_uid
)

etable(m_migrant_pre_A, m_migrant_pre_B, m_migrant_main,
       headers = c("Placebo 1997-2008 (women only)",
                   "Placebo 1997-2008 (sex as covariate)",
                   "Treatment 2008-2021"))
```


```{r}
#| label: fig-migrant-2x2
#| fig-cap: "2×2 DiD estimates for migrant"

did_row <- function(model, year_post, term = "treat_post") {
  tidy(model) |>
    filter(term == !!term) |>
    transmute(year = year_post, estimate, se = std.error)
}

migrant_2x2_df <- bind_rows(
  did_row(m_migrant_pre_B, 2008),
  did_row(m_migrant_main, 2021)
) |>
  mutate(
    period = factor(
      ifelse(year == 2008, "1997–2008 (placebo)", "2008–2021 (treatment)"),
      levels = c("1997–2008 (placebo)", "2008–2021 (treatment)")
    ),
    lo = estimate - 1.96 * se,
    hi = estimate + 1.96 * se
  )

ggplot(migrant_2x2_df, aes(x = period, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.2) +
  geom_point(size = 3) +
  labs(x = NULL, y = "DiD effect on the probability of migration",
       title = "Impact of protected areas on migration: 2×2 DiD") +
  theme_minimal(base_size = 13)
```


### Staggered DiD — static estimation

We use the Gardner (2022) two-stage DiD estimator (`did2s`) to leverage all available years with migration data (1997, 2008, 2018, 2021). This estimator is robust to heterogeneous treatment effects and staggered adoption.

```{r}
#| label: haz-staggered-static

# Prepare relative year bins (same scheme as main analysis)
dat_migration <- migration_pooled |>
  mutate(
    rel_year = DHSYEAR - STATUS_YR,
    rel_year_clipped = case_when(
      is.infinite(rel_year) ~ NA_real_,
      TRUE ~ pmax(pmin(rel_year, 10), -10)
    ),
    rel_year_2yr = case_when(
      is.na(rel_year_clipped) ~ NA_integer_,
      rel_year_clipped <= -9  ~ -5L,
      rel_year_clipped <= -7  ~ -4L,
      rel_year_clipped <= -5  ~ -3L,
      rel_year_clipped <= -3  ~ -2L,
      rel_year_clipped <= -1  ~ -1L,
      rel_year_clipped <=  1  ~  0L,
      rel_year_clipped <=  3  ~  1L,
      rel_year_clipped <=  5  ~  2L,
      rel_year_clipped <=  7  ~  3L,
      rel_year_clipped <=  9  ~  4L,
      TRUE                    ~  5L
    ),
    rel_year_2yr2 = case_when(
      is.na(rel_year_2yr) ~ "never",
      TRUE ~ as.character(rel_year_2yr)
    ),
    rel_year_2yr2 = factor(
      rel_year_2yr2,
      levels = c("-5", "-4", "-3", "-2", "-1", "0", "1", "2", "3", "4", "5", "never")
    ),
    treat_on = as.integer(treat == 1 & DHSYEAR >= STATUS_YR)
  )

did2s_static_migration <- did2s(
  data         = dat_migration ,
  yname        = "migrant",
  first_stage  = ~ sexe + education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(treat_on, ref = FALSE),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

modelsummary(list("did2s migration (static, pooled men + women)" = did2s_static_migration))
```


### Staggered DiD — event study

```{r}
#| label: migration-event-study

did2s_migration_es <- did2s(
  data         = dat_migration,
  yname        = "migrant",
  first_stage  = ~ sexe + education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_2yr2, ref = c("-1", "never")),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

etable(did2s_migration_es, headers = "did2s migration event study (2-year bins)")

```


```{r}
bin_label <- function(k) {
  dplyr::case_when(
    is.na(k) ~ NA_character_,
    k == "-5" ~ "<= -9",
    k == "-4" ~ "-8 to -7",
    k == "-3" ~ "-6 to -5",
    k == "-2" ~ "-4 to -3",
    k == "-1" ~ "-2 to -1",
    k == "0"  ~ "0 to 1",
    k == "1"  ~ "2 to 3",
    k == "2"  ~ "4 to 5",
    k == "3"  ~ "6 to 7",
    k == "4"  ~ "8 to 9",
    k == "5"  ~ ">= 10",
    TRUE      ~ as.character(k)
  )
}
 
event_study_df <- tidy(did2s_migration_es, conf.int = TRUE) |>
  filter(grepl("^rel_year_2yr2::", term)) |>
  mutate(
    k = sub("^rel_year_2yr2::", "", term),
    k = factor(k, levels = c("-5", "-4", "-3", "-2", "0", "1", "2", "3", "4", "5"))
  ) |>
  arrange(k)
 
ggplot(event_study_df, aes(k, estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  labs(x = "Years relative to protected area creation (2-year bins)",
       y = "Effect on the probability of migration",
       title = "Event study: migration around the creation of protected areas") +
  scale_x_discrete(labels = bin_label) +
  theme_minimal(base_size = 13)


```


## Migration by PA type - 2×2 DiD

We test whether effects differ between strict PAs (IUCN I–IV) and multi-use PAs (IUCN V–VI), following the same approach as the main analysis (H3). Given the sex-composition imbalance documented ealier (1997, 2011, 2013, 2016 are women-only waves), we use the women-only panel (`dat_migrant_femmes`)/ (`dat_migrant2_femmes`) as the base sample here, consistent with the main specification adopted for the pooled migration DiD. 

```{r}
#| label: migrant-iucn-2x2

dat_migrant3 <- dat_migrant_femmes |>
  mutate(
    treat_strict = as.integer(GROUP == "Treatment" & is_strict(IUCN_CAT)),
    treat_multi  = as.integer(GROUP == "Treatment" & is_multi(IUCN_CAT))
  )

# Placebo strict
pre_strict_migrant <- dat_migrant3 |>
  filter(DHSYEAR %in% c(1997, 2008), treat_strict == 1 | GROUP == "Control") |>
  mutate(post = as.integer(DHSYEAR == 2008))

m_pre_strict_migrant <- feols(
  migrant ~ treat_strict + post + treat_strict:post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = pre_strict_migrant, weights = ~w_all, cluster = ~cluster_uid
)

# Placebo multi
pre_multi_migrant <- dat_migrant3 |>
  filter(DHSYEAR %in% c(1997, 2008), treat_multi == 1 | GROUP == "Control") |>
  mutate(post = as.integer(DHSYEAR == 2008))

m_pre_multi_migrant <- feols(
  migrant ~ treat_multi + post + treat_multi:post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = pre_multi_migrant, weights = ~w_all, cluster = ~cluster_uid
)

# Treatment strict
main_strict_migrant <- dat_migrant3 |>
  filter(DHSYEAR %in% c(2008, 2021), treat_strict == 1 | GROUP == "Control") |>
  mutate(post = as.integer(DHSYEAR == 2021))

m_main_strict_migrant <- feols(
  migrant ~ treat_strict + post + treat_strict:post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = main_strict_migrant, weights = ~w_all, cluster = ~cluster_uid
)

# Treatment multi
main_multi_migrant <- dat_migrant3 |>
  filter(DHSYEAR %in% c(2008, 2021), treat_multi == 1 | GROUP == "Control") |>
  mutate(post = as.integer(DHSYEAR == 2021))

m_main_multi_migrant <- feols(
  migrant ~ treat_multi + post + treat_multi:post + education +
    spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220,
  data = main_multi_migrant, weights = ~w_all, cluster = ~cluster_uid
)

etable(m_pre_strict_migrant, m_main_strict_migrant, m_pre_multi_migrant, m_main_multi_migrant,
       headers = c("Strict plac.", "Strict treat.", "Multi plac.", "Multi treat."))
```


```{r}
#| label: fig-migrant-iucn-2x2
#| fig-cap: "2×2 DiD for migration by PA type"

migrant_iucn_df <- bind_rows(
  did_row(m_pre_strict_migrant, 2008, "treat_strict:post") |> mutate(type = "Strict"),
  did_row(m_main_strict_migrant, 2021, "treat_strict:post") |> mutate(type = "Strict"),
  did_row(m_pre_multi_migrant, 2008, "treat_multi:post") |> mutate(type = "Multi-use"),
  did_row(m_main_multi_migrant, 2021, "treat_multi:post") |> mutate(type = "Multi-use")
) |>
  mutate(
    period = factor(ifelse(year == 2008, "1997–2008", "2008–2021"),
                    levels = c("1997–2008", "2008–2021")),
    lo = estimate - 1.96 * se,
    hi = estimate + 1.96 * se
  )

ggplot(migrant_iucn_df, aes(x = period, y = estimate, colour = type)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 3, position = pd) +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.2, position = pd) +
  labs(x = NULL, y = "DiD effect on Pr(migration)",
       colour = "PA type",
       title = "Migration: 2×2 DiD by PA type") +
  theme_minimal(base_size = 13)
```


### Migration by PA type — staggered static

```{r}
#| label: migrant-iucn-staggered

d_strict_migrant <- dat_migrant2_femmes |>
  filter(GROUP == "Control" | (GROUP == "Treatment" & is_strict(IUCN_CAT))) |>
  mutate(treat_on = as.integer(treat == 1 & DHSYEAR >= STATUS_YR))

did2s_migrant_strict <- did2s(
  data         = d_strict_migrant,
  yname        = "migrant",
  first_stage  = ~ education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(treat_on, ref = FALSE),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

d_multi_migrant <- dat_migrant2_femmes |>
  filter(GROUP == "Control" | (GROUP == "Treatment" & is_multi(IUCN_CAT))) |>
  mutate(treat_on = as.integer(treat == 1 & DHSYEAR >= STATUS_YR))


did2s_migrant_multi <- did2s(
  data         = d_multi_migrant,
  yname        = "migrant",
  first_stage  = ~ education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(treat_on, ref = FALSE),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

modelsummary(list(
  "Migration Strict" = did2s_migrant_strict,
  "Migration Multi"  = did2s_migrant_multi,
  "Migration Both"   = did2s_migrant_static_femmes
))

```


### Migration by PA type — event study

```{r}
#| label: migrant-iucn-event-study

d_strict_migrant_es <- dat_migrant2_femmes |>
  filter(GROUP == "Control" | (GROUP == "Treatment" & is_strict(IUCN_CAT))) |>
  mutate(treat_on = as.integer(treat == 1 & DHSYEAR >= STATUS_YR))

did2s_migrant_es_strict <- did2s(
  data         = d_strict_migrant_es,
  yname        = "migrant",
  first_stage  = ~ education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_2yr2, ref = c("-1", "never")),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

d_multi_migrant_es <- dat_migrant2_femmes |>
  filter(GROUP == "Control" | (GROUP == "Treatment" & is_multi(IUCN_CAT))) |>
  mutate(treat_on = as.integer(treat == 1 & DHSYEAR >= STATUS_YR))

did2s_migrant_es_multi <- did2s(
  data         = d_multi_migrant_es,
  yname        = "migrant",
  first_stage  = ~ education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_2yr2, ref = c("-1", "never")),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

```



```{r}
# covariate on the full sample, which is not the right comparator here).
did2s_migrant_es_femmes <- did2s(
  data         = dat_migrant2_femmes,
  yname        = "migrant",
  first_stage  = ~ education + spei_wc_n_2 + spei_wc_n_1 + spei_wc_n + hv219 + hv220 | DHSYEAR,
  second_stage = ~ i(rel_year_2yr2, ref = c("-1", "never")),
  treatment    = "treat_on",
  cluster_var  = "cluster_uid",
  weights      = "w_all"
)

es_migrant_iucn <- bind_rows(
  tidy_es(did2s_migrant_es_strict, "Strict"),
  tidy_es(did2s_migrant_es_multi,  "Multi-use"),
  tidy_es(did2s_migrant_es_femmes, "Both")
)

ggplot(es_migrant_iucn, aes(k, estimate, colour = type)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(position = pd2, size = 2) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, position = pd2) +
  labs(x = "Years to treatment (2-year bins)",
       y = "Effect on Pr(migration)",
       colour = "PA type",
       title = "Event study: migration by PA type") +
  scale_x_discrete(labels = bin_label) +
  theme_minimal(base_size = 13)
```


## Summary of results

```{r}
summary_table <- bind_rows(
  extract_did2s(did2s_haz_static, "HAZ — All PAs"),
  extract_did2s(did2s_haz_strict, "HAZ — Strict"),
  extract_did2s(did2s_haz_multi, "HAZ — Multi-use"),
  extract_did2s(did2s_mort_static, "Mortality — All PAs"),
  extract_did2s(did2s_mort_strict, "Mortality — Strict"),
  extract_did2s(did2s_mort_multi, "Mortality — Multi-use"),
  extract_did2s(did2s_migrant_static_femmes, "Migration — All PAs (women)"),
  extract_did2s(did2s_migrant_strict, "Migration — Strict (women)"),
  extract_did2s(did2s_migrant_multi, "Migration — Multi-use (women)")
)
```


