---
title: "Audit — annual treatment timing, event-time bin support, and staggered-estimator feasibility"
subtitle: "Outcome-blind design audit for the staggered analysis (07/07b), review_v2"
date: "2026-10-07"
format:
  html:
    page-layout: full
    toc: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
out_dir <- "output/review_v2/event_time_support_audit"
img <- function(f, w = "100%") knitr::include_graphics(file.path("..", out_dir, f))
```

**Scope statement.** This audit evaluates design support only. No outcome variable was
loaded into an estimation routine or inspected during this audit. The historical 2-year
bin rule was recovered from code/manuscript predating the current outcome analysis; this
audit evaluates support for that pre-existing rule and does not alter bin boundaries based
on treatment-effect estimates.

## 1. Exact-date annual treatment time (`g_eff`)

Survey reference date = **June 1** of each survey year (frozen convention). For every
eventual-treated observation:

```r
g_eff = treatment_calendar_year + 1*(treatment_date > June 1 of that year)
```

This is not a new causal assumption: it is the annual representation of the frozen rule
`treated_now = survey_ref_date >= treatment_date`.

**Assertions (all pass, 6,087 treated-household observations):**

- `treated_now == as.integer(DHSYEAR >= g_eff)` holds for all treated observations.
- `GROUP == Control` ⟹ `treatment_date` NA and `g_eff == 0` (31,344 observations).
- One `(treatment_date, g_eff)` pair per treated PA (43 PAs).
- Cross-check: `floor((survey_ref_date − treatment_date)/365.25) == DHSYEAR − g_eff`
  for **all 6,087** treated observations. The date-consistent annual construction
  **exactly reproduces** the fractional event time recovered from the manuscript-era code
  (`07-estimation_staggered.qmd`), which itself superseded the older
  `DHSYEAR − treatment_year` construction.

**Annual cohorts (43 treated PAs):**

```{r}
#| label: cohort-table
cohort_pa <- tribble(
  ~treatment_date, ~n_PAs, ~g_eff,
  "2008-10-17", 31, 2009,
  "2008-12-02",  6, 2009,
  "2013-07-18",  1, 2014,
  "2015-01-01",  4, 2015,
  "2020-01-29",  2, 2020
)
cohort_pa
```

Four annual cohorts (2009, 2014, 2015, 2020) instead of the three survey-wave cohorts
(2011, 2016, 2021) used by the current 07b implementation.

## 2. Old vs date-consistent bin shifts

Relative time: old = `DHSYEAR − treatment_year`; date-consistent = `DHSYEAR − g_eff`.
The same historical 2-year bin rule is applied to both. 91.8% of treated observations
change their exact relative year (all PAs treated after June 1 shift by −1); 43.8% change
their 2-year bin; **all 43 PAs** change bin somewhere.

```{r}
#| label: shift-table
read_csv(file.path("..", out_dir, "bin_shift_observations.csv"), show_col_types = FALSE) |>
  rename(`old bin` = bin_old_lab, `date-consistent bin` = bin_eff_lab, `HH obs` = hh_obs)
```

The dominant shift: the old `0:1` bin (1,006 HH / 25 PAs — the 2008-cohort baseline wave)
moves into the reference bin `−2:−1`, because those PAs entered treatment **after** the
June 2008 interview. This matches the correction documented in the manuscript-era code.

```{r}
#| label: fig-shift
#| fig-cap: "Old integer bin vs date-consistent bin (households and distinct PAs)"
img("bin_shift_matrix.png")
```

## 3. Exact event-time support (date-consistent)

```{r}
#| label: fig-hist
#| fig-cap: "Exact relative-year histogram with the fixed historical bin overlay"
img("event_time_exact_histogram.png")
```

Only **16 distinct exact relative years** exist (cohort × wave arithmetic); the remaining
years are empty, which is precisely what the historical binning was designed to absorb.

```{r}
#| label: fig-binsup
#| fig-cap: "Household, cluster and PA support per historical bin"
img("event_time_binned_support.png")
```

Reading the three levels separately (rule: household mass alone is not credibility):

- `≤ −9`: 10 PAs / 12 clusters / 365 HH — spread over 4 cohorts, but driven by the
  1997 wave (rel −18, −17 from the 2015/2014 cohorts and −12 from the 2009 cohort).
- `−8:−7`, `−6:−5`, `−4:−3`: 1–3 PAs each — fragile at the PA level.
- `−2:−1` (reference bin): 27 PAs / 35 clusters / 1,064 HH — the strongest pre bin.
- `0:1`: 2 PAs only. `2:3`: 13 PAs / 2 cohorts. `4:5`: 16 PAs / 2 cohorts.
- `6:7`: 17 PAs / 3 cohorts. `8:9`: 22 PAs but **one cohort** (2009 @ 2018).
- `≥10`: 30 PAs but **one cohort** (2009 @ 2021).

## 4. Annual cohort support and cohort × wave structure

```{r}
#| label: fig-cohort
#| fig-cap: "PAs, clusters and post-treatment household observations per annual cohort"
img("annual_treatment_cohort_support.png")
```

```{r}
#| label: fig-cohwave
#| fig-cap: "Cohort × survey-wave support (post-treatment, treated side)"
img("annual_cohort_wave_support.png")
```

Critical support gaps visible in the full treated-side table (pre-treatment observations
included): the 2014 cohort has **no treated observations at 2011 or 2013**; the 2014 and
2015 cohorts are **entirely absent from the 2016 wave** despite being treated by then;
the 2020 cohort appears only at 2008, 2018 (pre) and 2021 (post).

```{r}
#| label: fig-collapse
#| fig-cap: "Annual cohorts collapsed to survey-wave cohorts in current 07b"
img("annual_vs_collapsed_cohorts.png")
```

```{r}
#| label: fig-bincomp
#| fig-cap: "PA support per bin, decomposed by annual cohort"
img("event_bin_cohort_composition.png")
```

## 5. C&S feasibility: does `did` accept annual `g_eff` between survey waves?

**YES.** Verified two ways against the installed **did 2.1.2** (R 4.5.1):

1. **Source inspection** (`pre_process_did`, `compute.att_gt`):
   - `gname` must be numeric but is never required to be a value of `tname`; groups are
     selected by `data$.G <- 1*(gname == glist[g])`.
   - `base_period = "universal"`: base = `tail(which(tlist < g), 1)` — the last observed
     wave strictly before `g_eff`. Default is `"varying"` (base = preceding wave), which
     is what 07b used.
   - RCS cells are skipped (NA) exactly when treated units are missing at `t` or at the
     base period, or controls are missing at either — a mechanical rule reproducible
     without outcomes.
   - Groups with `g > max(tlist)` are silently recoded never-treated (not applicable:
     max `g_eff` = 2020 < 2021).

2. **Synthetic toy test** (survey years matching ours; cohorts 2009/2014/2015/2020 between
   waves; synthetic outcome unrelated to the study): `att_gt(panel = FALSE,
   control_group = "nevertreated", est_method = "ipw")` produced all expected
   (g,t) cells under both base-period conventions.

### Estimability matrix (mechanical, no minimum-sample rules imposed)

```{r}
#| label: fig-estim
#| fig-cap: "ATT(g,t) estimability. Tile annotation on estimable cells: treated clusters/PAs at t : at base."
img("annual_attgt_estimability_matrix.png")
```

```{r}
#| label: estim-counts
cells <- read_csv(file.path("..", out_dir, "annual_attgt_cell_support.csv"), show_col_types = FALSE)
cells |>
  group_by(base_period, g_eff) |>
  summarise(`cells` = n(), `estimable` = sum(estimable), .groups = "drop") |>
  arrange(base_period, g_eff)
```

13 of 24 cells are mechanically estimable under the 07b default (`varying`); 13 of 28
under `universal` (which adds the g=2009 pre cell at 1997 and g=2015/2020 pre cells, but
loses g=2014 entirely). The g=2014 cohort has **no estimable cell under `universal`**
(its base wave 2013 contains no treated observations) and only 2 fragile cells under
`varying` (2008 pre: 1 PA; 2021 post: 1 PA at base).

### Outcome-blind overlap diagnostics (estimable post cells)

The exact `DRDID::std_ipw_did_rc` (v1.2.3) treatment-only machinery was replicated:
weighted logit propensity score on the pooled two-period comparison sample (cohort-g
units + never-treated controls; other cohorts excluded), PS capped at 1−1e-6, controls
with PS ≥ 0.995 trimmed, ATT odds weights. No outcomes involved.

```{r}
#| label: fig-overlap
#| fig-cap: "Max control PS, control ESS and post-IPW balance per estimable post cell"
img("annual_attgt_overlap_diagnostics.png")
```

- No control propensity score reaches the 0.995 trim in any cell (max observed 0.877,
  g=2015 at 2021 under `varying`).
- Control ESS is ample for g=2009 cells (≈ 4,400–4,500 at base); fragile but non-zero for
  the small cohorts (e.g., ESS ≈ 55 post / 188 base for g=2015 at 2021, `varying`).
- Post-IPW max |SMD| remains ≈ 0.2–0.5 for the g=2009 cells (same machinery as the frozen
  07b run, so inherited, not new).
- **The small-cohort cells show extreme covariate imbalance that is already present
  before IPW** (e.g., g=2020 at 2021: treated treecover mean 17,334 (SD 17) vs control
  20,318 (SD 8,263), raw |SMD| ≈ 245; g=2014: population 124 (SD 0.9) vs 29 (SD 25),
  raw |SMD| ≈ 96). The 2014/2015/2020 cohort PAs sit in landscapes far from the
  never-treated pool on the matching covariates, and their within-cohort dispersion is
  near zero. The frozen pooled matching balance (max |SMD| ≤ 0.10) does not carry over
  to these cohort-specific comparisons. This is a support problem, not a software one.
- One weighted logit showed near-separation ("fitted probabilities numerically 0 or 1")
  — flagged in `annual_attgt_overlap.csv`.

## 6. Aggregation of estimable cells into the historical bins

```{r}
#| label: fig-binnedcells
#| fig-cap: "Estimable ATT(g,t) cells, PAs and clusters pooled per historical bin"
img("binned_attgt_estimable_support.png")
```

```{r}
#| label: binned-table
read_csv(file.path("..", out_dir, "binned_attgt_support.csv"), show_col_types = FALSE) |>
  arrange(base_period, bin_lab) |>
  rename(bin = bin_lab, cells = n_cells, `# cohorts` = n_cohorts, PAs = n_PAs,
         clusters = n_clusters, `HH rows` = hh_rows)
```

**Key test result.** The historical bins do **not** meaningfully pool sparse annual
comparisons. Under the 07b default (`varying`), every bin except `6:7` draws on a
**single** annual cohort — and for all post-treatment bins that cohort is 2009. Under
`universal`, `≤ −9` pools 3 cohorts and `2:3`/`6:7` pool 2, but the g=2014 cohort
contributes nothing anywhere. The binned event study, if estimated from annual C&S cells,
is effectively a 2009-cohort event study at post horizons, with small-cohort contributions
concentrated in the `6:7` bin and scattered pre bins.

## 7. Estimator comparison (design feasibility only)

| | A. Callaway–Sant'Anna (`did::att_gt`) | B. Borusyak–Jaravel–Spiess imputation (`did2s`) | C. Old `did2s` implementation |
|---|---|---|---|
| Repeated cross-sections | Explicitly supported (`panel = FALSE`, verified in source and toy test) | `did2s` canonical use requires unit fixed effects from a panel; no RCS support claim in the installed package | Ran on RCS data, but see first-stage below |
| Treatment years between survey waves | **Yes** (annual `g_eff` accepted; base = last wave before `g_eff`) | Conceptually yes (imputation at observed periods), but targeting arbitrary binned event-time averages requires custom aggregation | Used date-exact fractional rel time, binned ±10 |
| Required stable-group structure | None beyond `gname` (treatment timing) | A persistent group/unit whose fixed effect can be estimated from untreated observations. DHS clusters (`hv001`) are **not** longitudinal; `region` is too coarse; PA-level FE is the only defensible candidate and changes the estimand to "within-PA" comparisons | **Only survey-wave FE** (`| DHSYEAR`) — no unit or group FE at all |
| Heterogeneity-robust | Yes (ATT(g,t) with aggregation) | Yes, if the group structure is valid | No — Gardner two-step with wave-FE-only first stage does not absorb time-invariant unit heterogeneity; binned second stage has no clean causal interpretation under effect heterogeneity |
| Clustering | Cluster bootstrap on wave×cluster (frozen 07b choice) | Unclear for RCS without units | `cluster_uid` = wave×cluster (not longitudinal) |
| Verdict | Mechanically feasible **today**; bins pool weakly across cohorts | Not adoptable without inventing a persistent-group structure; PA-FE variant would need methodological review | Legacy/exploratory; already flagged as not the reference estimator in the manuscript-era notes |

## 8. Decision

**Case B applies, with a sharp strong-cohort/weak-cohort structure.**

- Annual `g_eff` **is** accepted by `did` and exactly reproduces the manuscript-era
  fractional event time; the 2011/2016/2021 collapse was a design choice, not a software
  constraint (this rules out Case C).
- Mechanically, 13 ATT(g,t) cells are estimable, but the estimable set is dominated by
  the 2009 cohort (36 of 43 PAs); the 2014 cohort contributes at most 2 fragile
  single-PA cells, and the 2015/2020 cohorts rest on 1–3 PAs at base.
- The historical bins pool across cohorts almost nowhere; post-treatment bins are
  single-cohort by construction.
- Overlap: PS tails are fine (no trimming triggered), but the small-cohort cells show
  severe raw covariate displacement relative to the never-treated pool — the fragility is
  in *support*, not in the propensity-score machinery.
- BJS is **not** adoptable without a defensible stable-group structure; DHS clusters are
  not longitudinal units, and a PA-level FE variant would need separate methodological
  review. The old `did2s` implementation should not be revived merely for graphical
  convenience.

**Recommended next outcome-free methodological step.** Before any outcome is estimated,
decide — as an explicit design choice — between:

1. **Annual C&S with `varying` base (07b default) + historical bin aggregation**, reported
   with a per-cohort decomposition of every bin (this audit's exports make that
   transparent), acknowledging that post bins are 2009-cohort estimates; or
2. **Annual C&S with `universal` base**, which restores the g=2009 pre cell at 1997 and
   pools `≤ −9` across 3 cohorts, at the cost of losing g=2014 entirely — a trade-off
   that must be justified on design grounds, not on resulting estimates.

Either way: estimate C&S at annual cohorts first, aggregate into the fixed historical
bins second, and report bin-level cohort composition alongside every binned coefficient.

## 9. Files

- Script: `scripts/audit_annual_event_time_support.R`
- Outputs: `output/review_v2/event_time_support_audit/` (10 CSVs, 10 PNGs)
- No production file (`07-estimation_staggered.qmd`, `07b-estimation-staggered-did.qmd`)
  was modified.
